### 2019-02-14 README.Rmd
initial commit

---
output:
  github_document:
    html_preview: FALSE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# citesdb

```{r authors, echo = FALSE, results = 'asis'}
unclass(desc::desc_get_authors(here::here("DESCRIPTION"))) %>% 
  purrr::keep(~"aut" %in% .$role) %>% 
  purrr::map_chr(~paste(.$given, .$family)) %>% 
  paste(collapse = ", ") %>% 
  paste0("Authors: _", ., "_") %>% 
  cat()
```

[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)
[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
  [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ecohealthalliance/citesdb?branch=master&svg=true){data-external="1"}](https://ci.appveyor.com/project/ecohealthalliance/citesdb)
[![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
[![CRAN status](https://www.r-pkg.org/badges/version/monetpkg){data-external="1"}](https://cran.r-project.org/package=monetpkg)

Features:
-  Local caching of versions of database
-  Metadata in **[dataspice](https://github.com/ropenscilabs/dataspice)**/<schema.org> format
-  RStudio connection pane with browsable database
-  HTMLwidget-enhanced help files for browsing metadata?

## Installation

Install the **citesdb** package with this command:

```{r install_me, eval = FALSE}
source("https://install-github.me/ecohealthalliance/citesdb")
```

```{r message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
options(width=120)
```

## Usage

```{r example}
#library(citesdb)

# some example code
```

# Current Verison

```{r version}
packageVersion("monetpkg")
```

## Contributing

Want have feedback or want to contribute? Great! Please take a look at the [contributing guidelines](https://github.com/ecohealthalliance/citesdb/blob/master/.github/CONTRIBUTING.md) before filing an issue or pull request.

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/ecohealthalliance/citesdb/blob/master/.github/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2019-02-14 R/db.R

@@ -1,0 +1,16 @@
+#' @importFrom fs dir_info
+cites_db_size <- function(db = cites_db()) {
+  sum(fs::dir_info(cites_path(), recursive = TRUE)$size)
+}
+
+cites_db_export <- function(db) {
+
+}
+
+cites_db_import <- function(db)  {
+
+}
+
+cites_db_status <- function(db) {
+
+}

### 2019-02-14 R/download.R

@@ -1,0 +1,19 @@
+#' Title
+#'
+#' @param codes
+#' @param db
+#' @param new_only
+#' @param verbose
+#' @param temp_dir
+#' @param cleanup
+#'
+#' @return
+#' @export
+#'
+#' @examples
+cites_db_download <- function(codes = NULL, db = cites_db(), new_only = TRUE, verbose=TRUE,
+                            temp_dir = tempdir(), cleanup=TRUEs) {
+
+
+
+}

### 2019-02-14 R/onload.R

@@ -1,0 +1,84 @@
+#' @importFrom rappdirs user_data_dir
+cites_path <- function() {
+  sys_cites_path <- Sys.getenv("CITES_DB_HOME")
+  if (sys_cites_path == "") {
+    rappdirs::user_data_dir("citesdb")
+  }
+}
+
+#' @export
+cites_connect <- function(dbdir = cites_path()){
+  db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
+  if (inherits(db, "DBIConnection")) {
+    if (DBI::dbIsValid(db)) {
+      return(db)
+    }
+  }
+  dbname <- dbdir
+  dir.create(dbname, FALSE)
+  db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbname)
+  assign("cites_db", db, envir = cites_cache)
+  db
+}
+
+#' @export
+cites_disconnect <- function(env=cites_cache, shutdown = TRUE){
+  db <- mget("cites_db", envir = env, ifnotfound = NA)[[1]]
+  if (inherits(db, "DBIConnection"))
+    DBI::dbDisconnect(db, shutdown)
+  observer <- getOption("connectionObserver")
+  if (!is.null(observer))
+    observer$connectionClosed("MonetDB", "citesdb")
+}
+
+cites_cache <- new.env()
+reg.finalizer(cites_cache, cites_disconnect, onexit = TRUE)
+
+
+#' @export
+cites_pane <- function() {
+  observer <- getOption("connectionObserver")
+  if (!is.null(observer)) {
+    observer$connectionOpened(type = "MonetDB",
+                              host = "citesdb",
+                              displayName = "CITES",
+                              icon = system.file("img","eha_logo.png", package = "citesdb"),
+                              connectCode = "citesdb::cites_pane()",
+                              disconnect = citesdb::cites_disconnect,
+                              listObjectTypes = function() {
+                                list(
+                                  table = list(contains = "data")                                )
+                              },
+                              listObjects = function(type = "datasets") {
+                                tbls <- DBI::dbListTables(cites_connect())
+                                data.frame(
+                                  name = tbls,
+                                  type = rep("table", length(tbls)),
+                                  stringsAsFactors = FALSE
+                                )
+                              },
+                              listColumns = function(table) {
+                                res = DBI::dbSendQuery(cites_connect(), paste("SELECT * FROM", table, "LIMIT 1"))
[…]

### 2019-02-14 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(monetpkg)
+
+test_check("monetpkg")

### 2019-02-14 README.Rmd
some find/replace

@@
 )
+library(dplyr)
 ```
@@
 [![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
-[![CRAN status](https://www.r-pkg.org/badges/version/monetpkg){data-external="1"}](https://cran.r-project.org/package=monetpkg)
+[![CRAN status](https://www.r-pkg.org/badges/version/citesdb){data-external="1"}](https://cran.r-project.org/package=citesdb)
 
@@
 ```{r version}
-packageVersion("monetpkg")
+packageVersion("citesdb")
 ```

### 2019-02-14 R/download.R

@@ -17,3 +17,7 @@ cites_db_download <- function(codes = NULL, db = cites_db(), new_only = TRUE, ve
 
 
 }
+
+# Look at https://github.com/cboettig/taxadb/blob/master/R/td_create.R for this,
+# But use the MonetDB import functions (monetdb.read.csv) for speed.
+#

### 2019-02-14 tests/testthat.R

@@ -1,4 +1,4 @@
 library(testthat)
-library(monetpkg)
+library(citesdb)
 
-test_check("monetpkg")
+test_check("citesdb")

### 2019-02-15 README.Rmd
Author names

@@
   purrr::map_chr(~paste(.$given, .$family)) %>% 
-  paste(collapse = ", ") %>% 
+  glue::glue_collapse(sep = ", ", last = ", and ") %>% 
   paste0("Authors: _", ., "_") %>% 
@@
 ```{r version}
-packageVersion("citesdb")
+#packageVersion("citesdb")
 ```

### 2019-02-26 README.Rmd
functining package

@@
 
+**citesdb** provides access to the full history of CITES shipments for analysis.
+
 Features:
--  Local caching of versions of database
--  Metadata in **[dataspice](https://github.com/ropenscilabs/dataspice)**/<schema.org> format
+
+-  Local caching of database
 -  RStudio connection pane with browsable database
--  HTMLwidget-enhanced help files for browsing metadata?
+-  HTMLwidget-enhanced help files for browsing metadata
+
+While this repository is private you must have a GitHub Personal Access Token
+(`GITHUB_PAT`) installed in order to access data.  Instructions for setting this up are found
+[here](https://happygitwithr.com/github-pat.html).
 
@@
 
-## Usage
+## Getting the data
 
-```{r example}
-#library(citesdb)
+When you first load the package you will see a message like this:
 
-# some example code
+    library(citesdb)
+    #> Local CITES database empty or corrupt. Download with cites_db_download()
+
+Not to worry, just do as it says and run `cites_db_download()`.  This will
+fetch the most recent database from online.  Currently this is test data that
+is approximately 80MB.  It will expand to over 400MB in the database and during
+the import proccess up to 1GB of disk space may be used temporarily.
+
+Once you fetch the data you can connect to the database with the `cites_db()`
+command.  If you have the **dplyr** package installed, you can use the
+`cites_shipments()` command to load a remote `tibble()`.  You can use this
+to analyze CITES data without ever loading it into memory:
+
+```{r, warning = FALSE}
+library(citesdb)
+library(dplyr)
+start <- Sys.time()
+
+cites_shipments() %>%
+  group_by(year) %>%
+  summarize(number = n()) %>%
+  arrange(desc(year)) %>% 
+  collect()
+
+stop <- Sys.time()
 ```
 
-# Current Verison
+The back-end database, [MonetDB](https://monetdb.org) is very fast and powerful, making such
+analyses quite snappy even on such large data and normal desktops and laptops. Here's the
+timing of the above query on an ordinary laptop:
 
-```{r version}
-#packageVersion("citesdb")
+```{r}
+stop - start
 ```
 
-## Contributing
+If you are using RStudio interactively, loading the CITES package also brings
+up a browable pane in the "Connections" tab that lets you explore and preview
+the database, as well.
+
+### Contributing
 

### 2019-02-26 R/connect.R

@@ -6,8 +6,23 @@ cites_path <- function() {
   }
 }
 
+
+#' The local CITES database
+#'
+#' Returns a connecction to the local CITES database.  This is a DBI-complient
+#' [MonetDBLite::MonetDBLite]() database.
+#'
+#' @param dbdir The location of the database on disk.  Defaults to
+#' `citesdb` under [MonetDBLite::MonetDBLite], or `getOption("CITES_DB_DIR")`.
+#'
+#' @return A MonetDBLite DBI connection
+#' @importFrom DBI dbIsValid dbConnect
+#' @importFrom MonetDBLite MonetDBLite
 #' @export
-cites_connect <- function(dbdir = cites_path()){
+#'
+#' @examples
+#' cites_db()
+cites_db <- function(dbdir = cites_path()){
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
     if (DBI::dbIsValid(db)) {
@@ -16,23 +31,33 @@ cites_connect <- function(dbdir = cites_path()){
   }
   dbname <- dbdir
   dir.create(dbname, FALSE)
-  db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbname)
+  db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir)
   assign("cites_db", db, envir = cites_cache)
   db
 }
 
-cites_db <- cites_connect
-cites_trans <- function() {
-  if (!require(dplyr))
+
+#' CITES Shipment Data
+#'
+#' Returns a remote table with all CITES shipment data. Requires the dplyr and dbplyr packages.
+#' @return A dplyr remote tibble ([dplyr::tbl()])
+#' @export
+#'
+#' @examples
+#' cites_shipments()
+cites_shipments <- function() {
+  if (!cites_db_status(FALSE))
+    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
+  if (!suppressPackageStartupMessages(require(dplyr)))
     stop("Install the dplyr package to use convenience functions like cites_trans()")
-  dplyr::tbl(cites_connect(), "transactions")
+  dplyr::tbl(cites_db(), "shipments")
 }
 
 #' @export
 cites_disconnect <- function(env=cites_cache, shutdown = TRUE){
   db <- mget("cites_db", envir = env, ifnotfound = NA)[[1]]
[…]

### 2019-02-26 R/connection-pane.R

@@ -3,9 +3,9 @@ cites_pane <- function() {
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {
     observer$connectionOpened(
-      type = "MonetDB",
+      type = "CITESDB",
       host = "citesdb",
-      displayName = "CITES",
+      displayName = "CITES Transaction Tables",
       icon = system.file("img","eha_logo.png", package = "citesdb"),
       connectCode = "citesdb::cites_pane()",
       disconnect = citesdb::cites_disconnect,
@@ -14,7 +14,7 @@ cites_pane <- function() {
           table = list(contains = "data")                                )
       },
       listObjects = function(type = "datasets") {
-        tbls <- DBI::dbListTables(cites_connect())
+        tbls <- DBI::dbListTables(cites_db())
         data.frame(
           name = tbls,
           type = rep("table", length(tbls)),
@@ -22,20 +22,27 @@ cites_pane <- function() {
         )
       },
       listColumns = function(table) {
-        res = DBI::dbSendQuery(cites_connect(), paste("SELECT * FROM", table, "LIMIT 1"))
+        res = DBI::dbSendQuery(cites_db(), paste("SELECT * FROM", table, "LIMIT 1"))
         on.exit(DBI::dbClearResult(res))
         data.frame(name = res@env$info$names, type = res@env$info$types,
                    stringsAsFactors = FALSE)
       },
       previewObject = function(rowLimit, table) {
-        DBI::dbGetQuery(cites_connect(), paste("SELECT * FROM", table, "LIMIT", rowLimit))
+        DBI::dbGetQuery(cites_db(), paste("SELECT * FROM", table, "LIMIT", rowLimit))
       },
       actions = list(
-        boom = list(
-          icon = system.file("img","eha_logo.png", package = "citesdb"),
-          callback = function() cat("BOOM!")
+        status = list(
+          icon = system.file("img", "cites-logo.png", package = "citesdb"),
+          callback = cites_db_status
         )
       ),
-      connectionObject = cites_connect())
+      connectionObject = cites_db())
   }
 }
+
+update_cites_pane <- function() {
+  observer <- getOption("connectionObserver")
+  if (!is.null(observer))
+    observer$connectionUpdated("CITESDB","citesdb", "")
+}
+

### 2019-02-26 R/download.R

@@ -1,50 +1,107 @@
-#' Title
+#' Download the CITES database to your local computer
 #'
-#' @param codes
-#' @param db
-#' @param new_only
-#' @param verbose
-#' @param temp_dir
-#' @param cleanup
+#' This command downloads the CITES shipments database and populates a local
+#' database.  The download is large (>200MB), and the database will be at least
+#' twice that on disk. During import over 1GB of disk space may be used.
+#'
+#' @param tag What release tag of data to download. Defaults to the most recent
+#' @param destdir Where to download the compressed file.
+#' @param cleanup Whether to delete the compressed file after loading into the database.
+#' @param verbose Whether to display messages and download progress
 #'
 #' @return NULL
-#'
-#' @importFrom httr GET write_disk
+#' @export
+#' @importFrom DBI dbRemoveTable dbExistsTable dbCreateTable dbExecute dbWriteTable
+#' @importFrom R.utils gunzip
 #'
 #' @examples
-cites_db_download <- function(version = "latest",  dir = tempdir(),
-                              delete = TRUE, verbose = interactive()) {
+#' cites_db_download()
+cites_db_download <- function(tag = NULL,  destdir = tempdir(),
+                              cleanup = TRUE, verbose = interactive()) {
 
-  url <- citesdb_url(version)
-  cites_filename <- basename(url)
-  cites_file_path <- file.path(dir, cites_filename)
 
-  if (verbose) message("Downloading compressed data...\n")
-  response <- httr::GET(url, write_disk(cite_file_path), if (verbose) progress())
-  stop_for_status(response, "download CITES data")
+  if (verbose) message("Downloading data...\n")
+  zfile <- get_gh_release_file("ecohealthalliance/citesdb", tag_name = tag,
+                               destdir = destdir, verbose = verbose)
+  ver <- attr(zfile, "ver")
+  if (verbose) message("Decompressing and building local database...\n")
+  temp_tsv <- tempfile(fileext = ".tsv")
+  R.utils::gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
 
-  if (verbose) message("Building local database...\n")
-  DBI::dbRemoveTable(cites_db(), "transactions")
-  monet.read.csv(cites_db(), cites_file_path, header = TRUE, locked = TRUE)
[…]

### 2019-02-26 R/metadata.R

@@ -10,7 +10,7 @@
 #'
 #' \if{html}{
 #'   \Sexpr[echo=FALSE, results=rd, stage=build]{
-#'     if(cites:::is_js_ok()) {
+#'     if(citesdb:::is_js_ok()) {
 #'       mytext <- cites:::rd_datatable(cites::cites_codes())
 #'     } else {
 #'       mytext <- c('In RStudio help, this help file includes a searchable table of values if you install the DT package')
@@ -38,7 +38,7 @@ cites_codes <- function() {
 #'
 #' \if{html}{
 #'   \Sexpr[echo=FALSE, results=rd, stage=build]{
-#'     if (is_js_ok()) {
+#'     if (citesdb:::is_js_ok()) {
 #'       mytext <- cites:::rd_datatable(cites::cites_metadata())
 #'     } else {
 #'       mytext <- cites:::tabular(cites::cites_metadata())
@@ -66,7 +66,7 @@ cites_metadata <- function() {
 #'
 #' \if{html}{
 #'   \Sexpr[echo=FALSE, results=rd, stage=build]{
-#'     if (is_js_ok()) {
+#'     if (citesdb:::is_js_ok()) {
 #'       mytext <- cites:::rd_datatable(cites::cites_parties())
 #'     } else {
 #'       mytext <- cites:::tabular(cites::cites_parties())

### 2019-02-26 R/onattach.R

@@ -1,0 +1,54 @@
+.onAttach <- function(libname, pkgname) {
+
+  MonetDBLite::monetdblite_shutdown()
+  if (interactive() && Sys.getenv("RSTUDIO") == "1") {
+    cites_pane()
+  }
+  cites_db_status()
+}
+
+#' Remove the local CITES database
+#'
+#' Deletes all tables from the local database
+#'
+#' @return NULL
+#' @export
+#' @importFrom DBI dbListTables dbRemoveTable
+#'
+#' @examples
+#' cites_db_delete()
+cites_db_delete <- function() {
+  for (t in dbListTables(cites_db())) {
+    dbRemoveTable(cites_db(), t)
+  }
+  update_cites_pane()
+}
+
+
+#' Get the status of the current local CITES database
+#'
+#' @param verbose Whether to print a status message
+#'
+#' @return TRUE if the database exists, FALSE if it is not detected. (invisible)
+#' @export
+#' @importFrom DBI dbExistsTable
+#' @importFrom tools toTitleCase
+#' @examples
+#' cites_db_status()
+#'
+cites_db_status <- function(verbose = TRUE) {
+  if (dbExistsTable(cites_db(), "shipments") &&
+      dbExistsTable(cites_db(), "status")) {
+    status <- DBI::dbReadTable(cites_db(), "status")
+    status_msg <-
+      paste0("CITES database status:\n",
+             paste0(toTitleCase(gsub("_", " ", names(status))),
+                    ": ", as.matrix(status), collapse = "\n"))
+    out <- TRUE
+  } else {
+    status_msg <- "Local CITES database empty or corrupt. Download with cites_db_download()"
+    out <- FALSE
+  }
+  if (verbose) message(status_msg)
+  invisible(out)
+}

### 2019-02-26 README.Rmd
Tweak README

@@
 fetch the most recent database from online.  Currently this is test data that
-is approximately 80MB.  It will expand to over 400MB in the database and during
-the import proccess up to 1GB of disk space may be used temporarily.
+is approximately 80MB.  It will expand to over 400MB in the database. During
+up to 1GB of disk space may be used temporarily.
 
@@
 command.  If you have the **dplyr** package installed, you can use the
-`cites_shipments()` command to load a remote `tibble()`.  You can use this
-to analyze CITES data without ever loading it into memory:
+`cites_shipments()` command to load a remote `tibble` that is backed by the
+database but not loaded into R.  You can use this
+to analyze CITES data without ever loading it into memory, then gather your
+results with `collect()`:
 

### 2019-02-26 R/download.R
More ominous message

@@ -100,7 +100,7 @@ get_gh_release_file <- function(repo, tag_name = NULL, destdir = tempdir(),
   if (!length(release_obj)) stop("No release tagged \"", release, "\"")
 
   if (release_obj[[1]]$prerelease) {
-    message("This is pre-release data! It has not been validated.")
+    message("This is pre-release/sample data! It has not been cleaned or validated.")
   }
 
   download_url <- release_obj[[1]]$assets[[1]]$url

### 2019-03-16 README.Rmd
Update README
Use public CITES db release, bump to 0.1.0, +misc
- Change names in database
- Add convenience functions for metadata
- More completee README ready for public release
- Add metadata documentation from public DB
- Add SQL preview to connection pane

--- 2019-02-15 README.Rmd
@@
 
-Features:
--  Local caching of versions of database
--  Metadata in **[dataspice](https://github.com/ropenscilabs/dataspice)**/<schema.org> format
--  RStudio connection pane with browsable database
--  HTMLwidget-enhanced help files for browsing metadata?
+**citesdb** is a package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>.
+
+More details on the database can be found in the `?guidance` help file.
 
@@
 
-```{r example}
-#library(citesdb)
+### Getting the data
 
-# some example code
+When you first load the package you will see a message like this:
+
+    library(citesdb)
+    #> Local CITES database empty or corrupt. Download with cites_db_download()
+
+Not to worry, just do as it says and run `cites_db_download()`.  This will
+fetch the most recent database from online, an approximately 158 MB download.
+It will expand to over 1 GB in the local database. During
+the downloaad and database building up to 3.5 GB of disk space may be used temporarily.
+
+### Using the Database
+
+Once you fetch the data you can connect to the database with the `cites_db()`
+command.  You can use the
+`cites_shipments()` command to load a remote `tibble` that is backed by the
+database but not loaded into R.  You can use this
+to analyze CITES data without ever loading it into memory, then gather your
+results with `collect()`. For example:
+
+```{r, warning = FALSE}
+library(citesdb)
+library(dplyr)
+start <- Sys.time()
+
+cites_shipments() %>%
+  group_by(Year) %>%
+  summarize(number = n()) %>%
+  arrange(desc(Year)) %>%
+  collect()
+
+stop <- Sys.time()
 ```
 
-# Current Verison
+The back-end database, [MonetDB](https://monetdb.org) is very fast and powerful, making such
+analyses quite snappy even on such large data and normal desktops and laptops. Here's the
+timing of the above query, which procecsses over 20 million records, on an ordinary laptop:
 
-```{r version}
-#packageVersion("citesdb")
+```{r}
+stop - start
 ```
 
-## Contributing
+If you are using a recent version of RStudio interactively, loading the CITES package also brings
+up a browsable pane in the "Connections" tab that lets you explore and preview
+the database, as well as interact with it direcctly via SQL commands.
+
+### Metadata
+
+The database also contains tables of field metadata, codes used, and CITES countries. This information
+comes from ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf),
+on the CITES website.  Convenience functions `cites_metadata()`, `cites_codes()`, and `cites_parties()`
+access this information:
+
+```{r}
+head(cites_metadata())
+
+head(cites_codes())
+```
+
+More information on the release of shipment-level data can be found in the
+`?guidance` help file.
+
+### Contributing
 
@@
 By participating in this project you agree to abide by its terms.
+
+[![EcoHealth Allinace](https://github.com/ecohealthalliance/citesdb/blob/master/inst/img/eha-footer.png)](https://www.ecohealthalliance.org/)

### 2019-03-16 R/cites-guidance.R

@@ -1,0 +1,83 @@
+#' @title  CITES Trade Database shipment-level guidance
+#'
+#' @description _This text copied from the supplementary information file
+#'   accompanying the full database download from <https://trade.cites.org/>_
+#'
+#' @section Recommended citation:
+#'
+#'   UNEP-WCMC (Comps.) 2019. Full CITES Trade Database Download. Version
+#'   2019.2. CITES Secretariat, Geneva, Switzerland. Compiled by UNEP-WCMC,
+#'   Cambridge, UK. Available at: trade.cites.org.
+#'
+#' @section Background:
+#'
+#'   Under Article VIII of the Convention, Parties are required to provide
+#'   information regarding their trade in CITES-listed specimens through their
+#'   annual reports, and the Secretariat makes this information available
+#'   through the CITES Trade Database (trade.cites.org). This database currently
+#'   contains over 20 million records of international trade in CITES-listed
+#'   species. Parties recognise the importance of these reports as a tool for
+#'   monitoring the implementation of the Convention, assessing the
+#'   effectiveness of their wildlife management and trade policies, and to
+#'   enhance the detection of potentially harmful or illicit trade.
+#'
+#'   At the 70th meeting of the Standing Committee, Parties agreed that a full
+#'   non-aggregated version of the CITES Trade Database should be made available
+#'   and updated twice a year. These files represent the periodic release of the
+#'   CITES trade data in a shipment-by-shipment format with unique identifiers
+#'   replacing confidential permit numbers (see [SC70 Doc
+#'   26.2](https://cites.org/sites/default/files/eng/com/sc/70/E-SC70-26-02.pdf
+#'   ) and [SC70 Inf.
+#'   1](https://cites.org/sites/default/files/eng/com/sc/70/Inf/E-SC70-Inf-01.pdf)
+#'    for further background).
+#'
+#' @section Overview of data:
+#'
+#'   This zip file contains all trade records (including all historic data)
+#'   entered in the CITES Trade Database by 29 January 2019 and extracted at the
+#'   shipment level on 30 January 2019. This file is 2019.v2 and replaces
+#'   version 2019.v1 which contained some formatting anomalies and strengthens
+#'   security.
+#'
+#'   While the data provided through the search function on the [Web Portal of
+#'   the CITES Trade Database](https://trade.cites.org/) are aggregated, the
+#'   database contains non-aggregated data. The data provided in this download
+#'   is on a per-shipment basis i.e. it provides the relevant information about
[…]

### 2019-03-16 R/connect.R

@@ -45,14 +45,67 @@ cites_db <- function(dbdir = cites_path()) {
 #'
 #' @examples
 #' cites_shipments()
+#' @importFrom dplyr tbl
 cites_shipments <- function() {
-  if (!cites_db_status(FALSE)) {
+  if (!cites_status(FALSE)) {
     stop("Local CITES database empty or corrupt. Download with cites_db_download()")
   }
-  if (!suppressWarnings(suppressPackageStartupMessages(require(dplyr)))) {
-    stop("Install the dplyr package to use convenience functions like cites_trans()")
+  tbl(cites_db(), "cites_shipments")
+}
+
+#' Cites shipment metadata
+#'
+#' @description
+#'
+#' The CITES database also includes tables of column-levelel metadata and
+#' meanings of codes in colums, as well as a listing of CITES parties.
+#' convenience functions acccess these tables. As they are small, the functions
+#' collect the into R session memory, rather than returning a remote table.
+#'
+#' This information is drawn from
+#' ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf),
+#' from the CITES website. More information the the shipment-level data can be
+#' found in the [guidance] help file.
+#'
+#' @return A tibble of metadata
+#' @export
+#'
+#' @importFrom DBI dbReadTable
+#' @importFrom dplyr as_tibble
+#' @aliases metadata cites_metadata
+#' @examples
+#' cites_metadata()
+#' cites_codes()
+#' cites_shipments()
+#' 
+#' # For remote connections to these tables,
+#' # access the database directly:
+#' dplyr::tbl(cites_db(), "cites_metadata")
+#' dplyr::tbl(cites_db(), "cites_codes")
+#' dplyr::tbl(cites_db(), "cites_parties")
+cites_metadata <- function() {
+  if (!cites_status(FALSE)) {
+    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
   }
-  dplyr::tbl(cites_db(), "shipments")
+  as_tibble(dbReadTable(cites_db(), "cites_metadata"))
+}
+
+#' @export
+#' @rdname cites_metadata
+cites_codes <- function() {
+  if (!cites_status(FALSE)) {
+    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
+  }
+  as_tibble(dbReadTable(cites_db(), "cites_codes"))
+}
+
[…]

### 2019-03-16 R/connection-pane.R

@@ -1,3 +1,22 @@
+sql_action <- function() {
+  if (requireNamespace("rstudioapi", quietly = TRUE) &&
+    exists("documentNew", asNamespace("rstudioapi"))) {
+    contents <- paste(
+      "-- !preview conn=citesdb::cites_db()",
+      "",
+      "SELECT * FROM cites_shipments LIMIT 100",
+      "",
+      sep = "\n"
+    )
+
+    rstudioapi::documentNew(
+      text = contents, type = "sql",
+      position = rstudioapi::document_position(2, 25),
+      execute = FALSE
+    )
+  }
+}
+
 #' @export
 cites_pane <- function() {
   observer <- getOption("connectionObserver")
@@ -34,9 +53,13 @@ cites_pane <- function() {
         DBI::dbGetQuery(cites_db(), paste("SELECT * FROM", table, "LIMIT", rowLimit))
       },
       actions = list(
-        status = list(
+        Status = list(
           icon = system.file("img", "cites-logo.png", package = "citesdb"),
-          callback = cites_db_status
+          callback = cites_status
+        ),
+        SQL = list(
+          icon = system.file("img", "edit-sql.png", package = "citesdb"),
+          callback = sql_action
         )
       ),
       connectionObject = cites_db()

### 2019-03-16 R/download.R

@@ -28,7 +28,7 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
   temp_tsv <- tempfile(fileext = ".tsv")
   R.utils::gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
 
-  tblname <- "shipments"
+  tblname <- "cites_shipments"
   if (dbExistsTable(cites_db(), tblname)) {
     dbRemoveTable(cites_db(), tblname)
   }
@@ -47,24 +47,37 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
     )
   )
 
-  dbWriteTable(cites_db(), "status", make_status_table(version = ver),
+  dbWriteTable(cites_db(), "cites_status", make_status_table(version = ver),
     overwrite = TRUE
   )
 
   load_citesdb_metadata()
 
   file.remove(temp_tsv)
-  if (verbose) cites_db_status()
+  if (verbose) cites_status()
   update_cites_pane()
 }
 
 cites_field_types <- c(
-  id = "INTEGER", year = "INTEGER", appendix = "STRING", taxon = "STRING",
-  taxon_id = "INTEGER", class = "STRING", order = "STRING", family = "STRING",
-  genus = "STRING", reported_taxon = "STRING", reported_taxon_id = "INTEGER",
-  term = "STRING", quantity = "DOUBLE PRECISION", unit = "STRING", importer = "STRING",
-  exporter = "STRING", origin = "STRING", purpose = "STRING", source = "STRING",
-  reporter_type = "STRING"
+  Year = "INTEGER",
+  Appendix = "STRING",
+  Taxon = "STRING",
+  Class = "STRING",
+  Order = "STRING",
+  Family = "STRING",
+  Genus = "STRING",
+  Term = "STRING",
+  Quantity = "DOUBLE PRECISION",
+  Unit = "STRING",
+  Importer = "STRING",
+  Exporter = "STRING",
+  Origin = "STRING",
+  Purpose = "STRING",
+  Source = "STRING",
+  Reporter.type = "STRING",
+  Import.permit.RandomID = "STRING",
+  Export.permit.RandomID = "STRING",
+  Origin.permit.RandomID = "STRING"
 )
 
 #' @importFrom DBI dbGetQuery
@@ -74,7 +87,7 @@ make_status_table <- function(version) {
   data.frame(
     time_imported = Sys.time(),
     version = version,
[…]

### 2019-03-16 R/onattach.R

@@ -3,7 +3,7 @@
   if (interactive() && Sys.getenv("RSTUDIO") == "1") {
     cites_pane()
   }
-  if (interactive()) cites_db_status()
+  if (interactive()) cites_status()
 }
 
 #' Remove the local CITES database
@@ -33,11 +33,11 @@ cites_db_delete <- function() {
 #' @importFrom DBI dbExistsTable
 #' @importFrom tools toTitleCase
 #' @examples
-#' cites_db_status()
-cites_db_status <- function(verbose = TRUE) {
-  if (dbExistsTable(cites_db(), "shipments") &&
-    dbExistsTable(cites_db(), "status")) {
-    status <- DBI::dbReadTable(cites_db(), "status")
+#' cites_status()
+cites_status <- function(verbose = TRUE) {
+  if (dbExistsTable(cites_db(), "cites_shipments") &&
+    dbExistsTable(cites_db(), "cites_status")) {
+    status <- DBI::dbReadTable(cites_db(), "cites_status")
     status_msg <-
       paste0(
         "CITES database status:\n",
@@ -58,12 +58,16 @@ cites_db_status <- function(verbose = TRUE) {
 
 load_citesdb_metadata <- function() {
   tsvs <- list.files(system.file("extdata", package = "citesdb"),
-                     pattern = "\\.tsv$", full.names = TRUE)
+    pattern = "\\.tsv$", full.names = TRUE
+  )
   tblnames <- tools::file_path_sans_ext(basename(tsvs))
   for (i in seq_along(tsvs)) {
     suppressMessages(dbWriteTable(cites_db(), tblnames[i],
-                 read.table(tsvs[i], stringsAsFactors = FALSE, sep = "\t",
-                            header = TRUE,  quote = "\""),
-                 overwrite = TRUE))
+      read.table(tsvs[i],
+        stringsAsFactors = FALSE, sep = "\t",
+        header = TRUE, quote = "\""
+      ),
+      overwrite = TRUE
+    ))
   }
 }

### 2019-03-16 data-raw/clean_cites.R

@@ -1,39 +1,0 @@
-
-
-library(dplyr)
-library(readr)
-library(stringr)
-
-h <- here::here
-
-# ==============================================================================
-
-
-# Import CITES data sample
-
-cites.test <- read_csv("data-raw/cites_test_sample.csv")
-
-# Clean column names
-
-colnames(cites.test) <- colnames(cites.test) %>%
-  tolower() %>%
-  str_replace_all(., " ", "_")
-
-# Write compressed data to local disk
-
-write_tsv(
-  cites.test,
-  h("data-raw", "cites_test_data.tsv.gz")
-)
-
-# Release the compressed data
-
-# datastorr::github_release_info(
-#   "ecohealthalliance/citesdb",
-#   read = read_tsv,
-#   filename = h("data-raw", "cites_test_data.tsv.gz")
-# ) %>%
-# datastorr::github_release_create(
-#   description = "Release of CITES sample data (v0.0.0.9001)",
-#   target = "master", ignore_dirty = FALSE
-# )

### 2019-03-16 data-raw/process_trade_db.R

@@ -1,0 +1,71 @@
+library(httr)
+library(here)
+library(readr)
+library(fs)
+library(purrr)
+library(dplyr)
+GET(
+  "https://trade.cites.org/cites_trade/download_db",
+  write_disk(here("data-raw", "Trade_database.zip"), overwrite = TRUE)
+)
+unzip(here("data-raw", "Trade_database.zip"),
+  overwrite = TRUE,
+  exdir = here("data-raw"),
+)
+
+tdb <- map_dfr(
+  dir_ls(here("data-raw"), regexp = "^trade_db_\\d+\\.csv$"),
+  ~ read_csv(.,
+    col_types = cols(
+      Year = col_integer(),
+      Appendix = col_character(),
+      Taxon = col_character(),
+      Class = col_character(),
+      Order = col_character(),
+      Family = col_character(),
+      Genus = col_character(),
+      Term = col_character(),
+      Quantity = col_double(),
+      Unit = col_character(),
+      Importer = col_character(),
+      Exporter = col_character(),
+      Origin = col_character(),
+      Purpose = col_character(),
+      Source = col_character(),
+      Reporter.type = col_character(),
+      Import.permit.RandomID = col_character(),
+      Export.permit.RandomID = col_character(),
+      Origin.permit.RandomID = col_character()
+    )
+  )
+)
+
+# Removing invalid records to be fixed if possible in the future
+tdb <- tdb %>%
+  filter(!is.na(Year) & Year > 1970 & Year < 2020)
+
+# Arrange values; improves compressibility
+tdb <- tdb %>%
+  arrange(Year, Taxon, Order, Family, Genus, Term, Importer, Exporter, Appendix)
+
+# Write compressed data
+write_tsv(
+  tdb,
+  here("data-raw", "cites_trade_db.tsv.bz2")
+)
+
+# Clean up
+file_delete(
+  dir_ls(here("data-raw"), regexp = "(zip|csv|docx)$")
+)
+# Release the compressed data
+
+# datastorr::github_release_info(
+#   "ecohealthalliance/citesdb",
+#   read = read_tsv,
+#   filename = h("data-raw", "cites_trade_db.tsv.bz2")
+# ) %>%
+# datastorr::github_release_create(
+#   description = "Release of CITES shipment data (v2019.2)",
+#   target = "master", ignore_dirty = FALSE
+# )

### 2019-03-16 README.Rmd
Tweak README

@@
 [![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)
-[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
-  [![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/ecohealthalliance/citesdb?branch=master&svg=true){data-external="1"}](https://ci.appveyor.com/project/ecohealthalliance/citesdb)
-[![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
-[![CRAN status](https://www.r-pkg.org/badges/version/citesdb){data-external="1"}](https://cran.r-project.org/package=citesdb)
 
 **citesdb** is a package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>.
-
-More details on the database can be found in the `?guidance` help file.
+This data consists of over 20 million records of reported shipments of wildlife and wildlife
+products subject to oversight under the Convention on International Trade in Endangered Species).
 

### 2019-03-16 data-raw/process_trade_db.R

@@ -60,12 +60,12 @@ file_delete(
 )
 # Release the compressed data
 
-# datastorr::github_release_info(
-#   "ecohealthalliance/citesdb",
-#   read = read_tsv,
-#   filename = h("data-raw", "cites_trade_db.tsv.bz2")
-# ) %>%
-# datastorr::github_release_create(
-#   description = "Release of CITES shipment data (v2019.2)",
-#   target = "master", ignore_dirty = FALSE
-# )
+datastorr::github_release_info(
+  "ecohealthalliance/citesdb",
+  read = read_tsv,
+  filename = "cites_trade_db.tsv.bz2"
+) %>%
+datastorr::github_release_create(
+  description = "Release of CITES shipment data (v2019.2)",
+  target = "master", ignore_dirty = TRUE
+)

### 2019-03-16 README.Rmd
Fix typos

@@
 This data consists of over 20 million records of reported shipments of wildlife and wildlife
-products subject to oversight under the Convention on International Trade in Endangered Species).
+products subject to oversight under the Convention on International Trade in Endangered Species.
 
@@
 The back-end database, [MonetDB](https://monetdb.org) is very fast and powerful, making such
-analyses quite snappy even on such large data and normal desktops and laptops. Here's the
-timing of the above query, which procecsses over 20 million records, on an ordinary laptop:
+analyses quite snappy even on such large data using normal desktops and laptops. Here's the
+timing of the above query, which procecses over 20 million records, on an ordinary laptop:
 

### 2019-03-16 README.Rmd
Misc cleanup

@@
 This data consists of over 20 million records of reported shipments of wildlife and wildlife
-products subject to oversight under the Convention on International Trade in Endangered Species.
+products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org)
 
@@
 It will expand to over 1 GB in the local database. During
-the downloaad and database building up to 3.5 GB of disk space may be used temporarily.
+the download and database building up to 3.5 GB of disk space may be used temporarily.
 
@@
 analyses quite snappy even on such large data using normal desktops and laptops. Here's the
-timing of the above query, which procecses over 20 million records, on an ordinary laptop:
+timing of the above query, which processes over 20 million records, on an ordinary laptop:
 
@@
 up a browsable pane in the "Connections" tab that lets you explore and preview
-the database, as well as interact with it direcctly via SQL commands.
+the database, as well as interact with it directly via SQL commands.
 
@@
 
-[![EcoHealth Allinace](https://github.com/ecohealthalliance/citesdb/blob/master/inst/img/eha-footer.png)](https://www.ecohealthalliance.org/)
+[![Created by EcoHealth Alliance](https://github.com/ecohealthalliance/citesdb/blob/master/inst/img/eha-footer.png)](https://www.ecohealthalliance.org/)

### 2019-03-16 R/citesdb-package.R

@@ -3,10 +3,10 @@
 #' **citesdb** provides over 40 years of trade data from the [Convention on
 #' International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
 #' These data on imports and exports of CITES annex species are provided by
-#' CITES membere countries to the CITES secratariat.
+#' CITES member countries to the CITES secretariat.
 #'
-#' This package allows you to download the full database of histric transactions
-#' to anayze locally on your computer. Because of the large size of this data,
+#' This package allows you to download the full database of historic transactions
+#' to analyze locally on your computer. Because of the large size of this data,
 #' the **citesdb** package stores it as an on-disk [MonetDB](https://www.monetdb.org/)
 #' that can be queried without loading into RAM.
 #'

### 2019-03-16 R/connect.R

@@ -9,8 +9,8 @@ cites_path <- function() {
 
 #' The local CITES database
 #'
-#' Returns a connecction to the local CITES database.  This is a DBI-complient
-#' [MonetDBLite::MonetDBLite]() database.
+#' Returns a connection to the local CITES database.  This is a DBI-compliant
+#' [MonetDBLite::MonetDBLite]() database connection.
 #'
 #' @param dbdir The location of the database on disk.  Defaults to
 #' `citesdb` under [MonetDBLite::MonetDBLite], or `getOption("CITES_DB_DIR")`.
@@ -57,9 +57,9 @@ cites_shipments <- function() {
 #'
 #' @description
 #'
-#' The CITES database also includes tables of column-levelel metadata and
-#' meanings of codes in colums, as well as a listing of CITES parties.
-#' convenience functions acccess these tables. As they are small, the functions
+#' The CITES database also includes tables of column-level metadata and
+#' meanings of codes in columns, as well as a listing of CITES parties.
+#' convenience functions access these tables. As they are small, the functions
 #' collect the into R session memory, rather than returning a remote table.
 #'
 #' This information is drawn from
@@ -77,7 +77,7 @@ cites_shipments <- function() {
 #' cites_metadata()
 #' cites_codes()
 #' cites_shipments()
-#' 
+#'
 #' # For remote connections to these tables,
 #' # access the database directly:
 #' dplyr::tbl(cites_db(), "cites_metadata")
@@ -108,12 +108,12 @@ cites_parties <- function() {
   as_tibble(dbReadTable(cites_db(), "cites_metadata"))
 }
 
-#' Discconnect from the CITES database
+#' Disconnect from the CITES database
 #'
 #' A utility function for disconnecting from the database.
 #'
 #' @examples
-#' cites_disconnecct()
+#' cites_disconnect()
 #' @export
 cites_disconnect <- function(env = cites_cache, shutdown = TRUE) {
   db <- mget("cites_db", envir = env, ifnotfound = NA)[[1]]

### 2019-03-16 README.Rmd
Misc cleanup

@@
 This data consists of over 20 million records of reported shipments of wildlife and wildlife
-products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org)
+products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
 
@@
 
-### Using the Database
+### Using the database
 

### 2019-03-16 README.Rmd
Misc cleanup

@@
 
-The back-end database, [MonetDB](https://monetdb.org) is very fast and powerful, making such
+The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making such
 analyses quite snappy even on such large data using normal desktops and laptops. Here's the

### 2019-03-16 R/connect.R
Pass R CMD check

@@ -44,6 +44,7 @@ cites_db <- function(dbdir = cites_path()) {
 #' @export
 #'
 #' @examples
+#' if (cites_status())
 #' cites_shipments()
 #' @importFrom dplyr tbl
 cites_shipments <- function() {
@@ -74,15 +75,16 @@ cites_shipments <- function() {
 #' @importFrom dplyr as_tibble
 #' @aliases metadata cites_metadata
 #' @examples
-#' cites_metadata()
-#' cites_codes()
-#' cites_shipments()
+#' if (cites_status()) {
+#'   cites_metadata()
+#'   cites_codes()
+#'   cites_shipments()
 #'
-#' # For remote connections to these tables,
-#' # access the database directly:
-#' dplyr::tbl(cites_db(), "cites_metadata")
-#' dplyr::tbl(cites_db(), "cites_codes")
-#' dplyr::tbl(cites_db(), "cites_parties")
+#'   # For remote connections to these tables,
+#'   # access the database directly:
+#'   dplyr::tbl(cites_db(), "cites_metadata")
+#'   dplyr::tbl(cites_db(), "cites_codes")
+#' }
 cites_metadata <- function() {
   if (!cites_status(FALSE)) {
     stop("Local CITES database empty or corrupt. Download with cites_db_download()")
@@ -115,8 +117,12 @@ cites_parties <- function() {
 #' @examples
 #' cites_disconnect()
 #' @export
-cites_disconnect <- function(env = cites_cache, shutdown = TRUE) {
-  db <- mget("cites_db", envir = env, ifnotfound = NA)[[1]]
+#'
+cites_disconnect <- function() {
+  cites_disconnect_()
+}
+cites_disconnect_ <- function(environment = cites_cache) {
+  db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
     MonetDBLite::monetdblite_shutdown()
   }
@@ -126,12 +132,5 @@ cites_disconnect <- function(env = cites_cache, shutdown = TRUE) {
   }
 }
 
-cites_clean <- function(db = db_connect()) {
-  tables <- DBI::dbListTables(db)
-  drop <- tables[ !grepl("_", tables) ]
-  lapply(drop, function(x) DBI::dbRemoveTable(db, x))
-  invisible(TRUE)
-}
-
 cites_cache <- new.env()
[…]

### 2019-03-16 R/connection-pane.R

@@ -17,7 +17,17 @@ sql_action <- function() {
   }
 }
 
+#' Open CITES database connecction pane in RStudio
+#'
+#' This function launches the RStudio "Connection" pane to interactively
+#' explore the database.
+#'
+#' @return NULL
 #' @export
+#'
+#' @examples
+#' if (!is.null(getOption("connectionObserver")))
+#'   cites_pane()
 cites_pane <- function() {
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {

### 2019-03-16 R/download.R

@@ -15,7 +15,11 @@
 #' @importFrom R.utils gunzip
 #'
 #' @examples
+#' \donttest{
+#' \dontrun{
 #' cites_db_download()
+#' }
+#' }
 cites_db_download <- function(tag = NULL, destdir = tempdir(),
                               cleanup = TRUE, verbose = interactive()) {
   if (verbose) message("Downloading data...\n")
@@ -109,7 +113,7 @@ get_gh_release_file <- function(repo, tag_name = NULL, destdir = tempdir(),
     release_obj <- purrr::keep(releases, function(x) x$tag_name == tag_name)
   }
 
-  if (!length(release_obj)) stop("No release tagged \"", release, "\"")
+  if (!length(release_obj)) stop("No release tagged \"", tag_name, "\"")
 
   if (release_obj[[1]]$prerelease) {
     message("This is pre-release/sample data! It has not been cleaned or validated.")

### 2019-03-16 R/onattach.R

@@ -15,7 +15,11 @@
 #' @importFrom DBI dbListTables dbRemoveTable
 #'
 #' @examples
+#' \donttest{
+#' \dontrun{
 #' cites_db_delete()
+#' }
+#' }
 cites_db_delete <- function() {
   for (t in dbListTables(cites_db())) {
     dbRemoveTable(cites_db(), t)
@@ -56,6 +60,7 @@ cites_status <- function(verbose = TRUE) {
 }
 
 
+#' @importFrom utils read.table
 load_citesdb_metadata <- function() {
   tsvs <- list.files(system.file("extdata", package = "citesdb"),
     pattern = "\\.tsv$", full.names = TRUE

### 2019-03-16 R/rd-helpers.R

@@ -1,58 +1,0 @@
-# Modified from https://cran.r-project.org/web/packages/roxygen2/vignettes/formatting.html#tables
-
-tabular <- function(df, col_names = TRUE, ...) {
-  stopifnot(is.data.frame(df))
-
-  align <- function(x) if (is.numeric(x)) "r" else "l"
-  col_align <- vapply(df, align, character(1))
-
-  cols <- lapply(df, format, ...)
-  contents <- do.call(
-    "paste",
-    c(cols, list(sep = " \\tab ", collapse = "\\cr\n  "))
-  )
-
-  if (col_names) {
-    header <- paste0("\\bold{", colnames(df), "}", collapse = " \\tab")
-    contents <- paste0(header, "\\cr\n  ", contents)
-  }
-
-  paste(
-    "\\tabular{", paste(col_align, collapse = ""), "}{\n  ",
-    contents, "\n}\n",
-    sep = ""
-  )
-}
-
-#' @noRd
-rd_datatable <- function(df, width = "100%", ...) {
-  wrap_widget(DT::datatable(df, width = width, ...))
-}
-
-#' @noRd
-wrap_widget <- function(widget) {
-  tmp <- tempfile(fileext = ".html")
-  htmlwidgets::saveWidget(widget, tmp)
-  widg <- paste(
-    grep("^</?(!DOCTYPE|meta|body|html|head|title)",
-      readLines(tmp),
-      value = TRUE, invert = TRUE
-    ),
-    collapse = "\n"
-  )
-  paste("\\out{", escape_rd(widg), "}\n", sep = "\n")
-}
-
-#' @noRd
-escape_rd <- function(x) {
-  x <- gsub("\\", "\\\\", x, fixed = TRUE)
-  x <- gsub("%", "\\%", x, fixed = TRUE)
-  x <- gsub("{", "\\{", x, fixed = TRUE)
-  x <- gsub("}", "\\}", x, fixed = TRUE)
-  x
-}
-
-is_js_ok <- function() {
-  in_pkgdown <- any(grepl("as_html.tag_Sexpr", sapply(sys.calls(), function(a) paste(deparse(a), collapse = "\n"))))
-  !(in_pkgdown) && require(DT)
-}

### 2019-03-17 README.Rmd
Add comment about raw data source
Pass goodpractice except for covr, set up CI

@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
+[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=svg){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
 [![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)
@@
 
+If you don't need any of the bells and whistles of this package, you can download
+the raw data as a single compressed TSV from the [releases page](https://github.com/ecohealthalliance/citesdb/releases),
+or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.
+
 ### Metadata

### 2019-03-17 R/connect.R

@@ -6,14 +6,19 @@ cites_path <- function() {
   }
 }
 
+check_status <- function() {
+  if (!cites_status(FALSE)) {
+    stop("Local CITES database empty or corrupt. Download with cites_db_download()")  #nolint
+  }
+}
 
 #' The local CITES database
 #'
 #' Returns a connection to the local CITES database.  This is a DBI-compliant
-#' [MonetDBLite::MonetDBLite]() database connection.
+#' [MonetDBLite::MonetDBLite()] database connection.
 #'
 #' @param dbdir The location of the database on disk.  Defaults to
-#' `citesdb` under [MonetDBLite::MonetDBLite], or `getOption("CITES_DB_DIR")`.
+#' `citesdb` under [MonetDBLite::MonetDBLite()], or `getOption("CITES_DB_DIR")`.
 #'
 #' @return A MonetDBLite DBI connection
 #' @importFrom DBI dbIsValid dbConnect
@@ -48,9 +53,7 @@ cites_db <- function(dbdir = cites_path()) {
 #' cites_shipments()
 #' @importFrom dplyr tbl
 cites_shipments <- function() {
-  if (!cites_status(FALSE)) {
-    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
-  }
+  check_status()
   tbl(cites_db(), "cites_shipments")
 }
 
@@ -86,27 +89,21 @@ cites_shipments <- function() {
 #'   dplyr::tbl(cites_db(), "cites_codes")
 #' }
 cites_metadata <- function() {
-  if (!cites_status(FALSE)) {
-    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
-  }
+  check_status
   as_tibble(dbReadTable(cites_db(), "cites_metadata"))
 }
 
 #' @export
 #' @rdname cites_metadata
 cites_codes <- function() {
-  if (!cites_status(FALSE)) {
-    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
-  }
+  check_status()
   as_tibble(dbReadTable(cites_db(), "cites_codes"))
 }
 
 #' @export
 #' @rdname cites_metadata
 cites_parties <- function() {
-  if (!cites_status(FALSE)) {
-    stop("Local CITES database empty or corrupt. Download with cites_db_download()")
-  }
+  check_status()
   as_tibble(dbReadTable(cites_db(), "cites_metadata"))
 }
 
[…]

### 2019-03-17 R/connection-pane.R

@@ -11,7 +11,7 @@ sql_action <- function() {
 
     rstudioapi::documentNew(
       text = contents, type = "sql",
-      position = rstudioapi::document_position(2, 25),
+      position = rstudioapi::document_position(2, 40),
       execute = FALSE
     )
   }
@@ -52,15 +52,17 @@ cites_pane <- function() {
         )
       },
       listColumns = function(table) {
-        res <- DBI::dbSendQuery(cites_db(), paste("SELECT * FROM", table, "LIMIT 1"))
+        res <- DBI::dbSendQuery(cites_db(),
+                                paste("SELECT * FROM", table, "LIMIT 1"))
         on.exit(DBI::dbClearResult(res))
         data.frame(
           name = res@env$info$names, type = res@env$info$types,
           stringsAsFactors = FALSE
         )
       },
-      previewObject = function(rowLimit, table) {
-        DBI::dbGetQuery(cites_db(), paste("SELECT * FROM", table, "LIMIT", rowLimit))
+      previewObject = function(rowLimit, table) {  #nolint
+        DBI::dbGetQuery(cites_db(),
+                        paste("SELECT * FROM", table, "LIMIT", rowLimit))
       },
       actions = list(
         Status = list(

### 2019-03-17 R/download.R

@@ -30,7 +30,7 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
   ver <- attr(zfile, "ver")
   if (verbose) message("Decompressing and building local database...\n")
   temp_tsv <- tempfile(fileext = ".tsv")
-  R.utils::gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
+  gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
 
   tblname <- "cites_shipments"
   if (dbExistsTable(cites_db(), tblname)) {
@@ -86,24 +86,30 @@ cites_field_types <- c(
 
 #' @importFrom DBI dbGetQuery
 make_status_table <- function(version) {
-  sz <- sum(file.info(list.files(cites_path(), all.files = TRUE, recursive = TRUE, full.names = TRUE))$size)
+  sz <- sum(file.info(list.files(cites_path(),
+                                 all.files = TRUE,
+                                 recursive = TRUE,
+                                 full.names = TRUE))$size)
   class(sz) <- "object_size"
   data.frame(
     time_imported = Sys.time(),
     version = version,
-    number_of_records = formatC(DBI::dbGetQuery(cites_db(), "SELECT COUNT(*) FROM cites_shipments;")[[1]], format = "d", big.mark = ","),
+    number_of_records = formatC(
+      DBI::dbGetQuery(cites_db(),
+                      "SELECT COUNT(*) FROM cites_shipments;")[[1]],
+      format = "d", big.mark = ","),
     size_on_disk = format(sz, "auto")
   )
 }
 
-#' @import httr
+#' @importFrom httr GET stop_for_status content accept write_disk progress
 #' @importFrom purrr keep
 get_gh_release_file <- function(repo, tag_name = NULL, destdir = tempdir(),
                                 overwrite = TRUE, verbose = interactive()) {
   releases <- GET(
     paste0("https://api.github.com/repos/", repo, "/releases")
   )
-  httr::stop_for_status(releases, "finding releases")
+  stop_for_status(releases, "finding releases")
 
   releases <- content(releases)
 
@@ -116,7 +122,7 @@ get_gh_release_file <- function(repo, tag_name = NULL, destdir = tempdir(),
   if (!length(release_obj)) stop("No release tagged \"", tag_name, "\"")
 
   if (release_obj[[1]]$prerelease) {
-    message("This is pre-release/sample data! It has not been cleaned or validated.")
[…]

### 2019-03-17 R/metadata.R

@@ -1,86 +1,0 @@
-# CITES Code Values
-#
-# This function returns a data frame with descriptions of all the code values
-# used in [cites_data()].  This is useful for lookup
-# as well as merging with the data for more descriptive summaries.
-#
-# These values are drawn from
-# ["A guide to using the CITES Trade Database"](https://github.com/ecohealthalliance/cites/tree/master/inst/extdata),
-# from the CITES website.
-#
-# \if{html}{
-#   \Sexpr[echo=FALSE, results=rd, stage=build]{
-#     if(citesdb:::is_js_ok()) {
-#       mytext <- citesdb:::rd_datatable(citesdb::cites_codes())
-#     } else {
-#       mytext <- c('In RStudio help, this help file includes a searchable table of values if you install the DT package')
-#     }
-#     mytext
-#   }
-# }
-#
-# #' \if{text,latex}{The HTML version of this help file includes a searchable table of the CITES codes.}
-# #'
-# #' @return A tibble with fields and descriptions
-# #' @seealso [cites_metadata()] [cites_data()]
-# #' @export
-# cites_codes <- function() {
-#   cites_codes_
-# }
-
-# CITES Field Descriptions
-#
-# This function returns a data frame field descriptions for [cites_data()].
-#
-# This information is taken from
-# ["A guide to using the CITES Trade Database"](https://github.com/ecohealthalliance/cites/tree/master/inst/extdata),
-# from the CITES websites.
-#
-# \if{html}{
-#   \Sexpr[echo=FALSE, results=rd, stage=build]{
-#     if (citesdb:::is_js_ok()) {
-#       mytext <- citesdb:::rd_datatable(citesdb::cites_metadata())
-#     } else {
-#       mytext <- citesdb:::tabular(citesdb::cites_metadata())
-#     }
-#     mytext
-#   }
-# }
-#
-# #' \if{text,latex}{ \Sexpr[echo=FALSE, results=rd, stage=build]{citesdb:::tabular(citesdb::cites_metadata())}}
-# #'
-# #' @return A tibble with field, code, and code description
-# #' @aliases metadata
-# #' @seealso [cites_codes()] [cites_data()] [cites_parties()]
-# #' @export
-# # cites_metadata <- function() {
-# #   cites_metadata_
-# # }
-#
-#
-
-# Parties to the CITES treaty.
-#
[…]

### 2019-03-17 R/onattach.R

@@ -52,7 +52,7 @@ cites_status <- function(verbose = TRUE) {
       )
     out <- TRUE
   } else {
-    status_msg <- "Local CITES database empty or corrupt. Download with cites_db_download()"
+    status_msg <- "Local CITES database empty or corrupt. Download with cites_db_download()" #nolint
     out <- FALSE
   }
   if (verbose) message(status_msg)

### 2019-03-17 README.Rmd
Add codecov badge

@@
 [![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=svg){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
 [![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)

### 2019-03-17 README.Rmd
Use sheild style CI badge [ci skip]

--- 2019-03-17 README.Rmd
@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
-[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=svg){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=sheild){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
 [![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)

### 2019-03-17 README.Rmd
Fix sheild. revert to r-hub check at end

--- 2019-03-17 README.Rmd
@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
-[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=svg){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=shield){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
 [![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)

### 2019-03-18 tests/testthat/test-04-data.R
Add test for updated data

@@ -1,0 +1,10 @@
+context("Data")
+
+test_that("CITES data source hasn't changed", {
+  skip_on_cran()
+  httr::GET("https://trade.cites.org/cites_trade/download_db",
+            httr::write_disk("cites_trade_db.zip"))
+  expect_equivalent(tools::md5sum("cites_trade_db.zip"),
+                    "3d774b109c3ebf3594cf0fd4c20c1d1b")
+  unlink("cites_trade_db.zip")
+})

### 2019-03-18 R/connect.R
Add informative error and test for database lock

@@ -10,7 +10,7 @@ cites_path <- function() {
 
 check_status <- function() {
   if (!cites_status(FALSE)) {
-    stop("Local CITES database empty or corrupt. Download with cites_db_download()")  #nolint
+    stop("Local CITES database empty or corrupt. Download with cites_db_download()") # nolint
   }
 }
 
@@ -38,7 +38,24 @@ cites_db <- function(dbdir = cites_path()) {
   }
   dbname <- dbdir
   dir.create(dbname, FALSE)
-  db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir)
+
+  tryCatch(
+    db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir),
+    error = function(e) {
+      if (grepl("Database lock", e)) {
+        stop(paste(
+          "Local taxadb database is locked by another R session.\n",
+          "Try closing or running cites_disconect() in that session."
+        ),
+        call. = FALSE
+        )
+      } else {
+        stop(e)
+      }
+    },
+    finally = NULL
+  )
+
   assign("cites_db", db, envir = cites_cache)
   db
 }
@@ -51,8 +68,9 @@ cites_db <- function(dbdir = cites_path()) {
 #' @export
 #'
 #' @examples
-#' if (cites_status())
-#' cites_shipments()
+#' if (cites_status()) {
+#'   cites_shipments()
+#' }
 #' @importFrom dplyr tbl
 cites_shipments <- function() {
   check_status()
@@ -121,7 +139,7 @@ cites_parties <- function() {
 cites_disconnect <- function() {
   cites_disconnect_()
 }
-cites_disconnect_ <- function(environment = cites_cache) { #nolint
+cites_disconnect_ <- function(environment = cites_cache) { # nolint
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
     MonetDBLite::monetdblite_shutdown()

### 2019-03-18 tests/testthat/test-03-connection.R

@@ -1,10 +1,23 @@
 olddir <- Sys.getenv("CITES_DB_DIR")
 Sys.setenv(CITES_DB_DIR = normalizePath(file.path(getwd(), "localdb"),
-  mustWork = FALSE
+                                        mustWork = FALSE
 ))
 
+context("Connection")
 
-context("Shutdown")
+test_that("Database lock gives informative error", {
+  skip_on_cran()
+  skip_if_not(cites_status())
+  cites_db()
+  expect_error({
+    callr::r(function() {
+      options(CITES_DB_DIR = "localdb")
+      citesdb::cites_db()
+    })
+  },
+  "Local taxadb database is locked by another R session"
+  )
+})
 
 test_that("Disconnetion works", {
   skip_on_cran()
@@ -12,20 +25,20 @@ test_that("Disconnetion works", {
 
   cites_disconnect()
   expect_error({
-      success <- callr::r(function() {
-        options(CITES_DB_DIR = "localdb")
-        con <- DBI::dbConnect(
-          MonetDBLite::MonetDBLite(),
-          getOption("CITES_DB_DIR")
-        )
-        out <- inherits(con, "MonetDBEmbeddedConnection")
-        citesdb::cites_disconnect()
-        options(CITES_DB_DIR = NULL)
-        return(out)
-      })
-      stopifnot(success)
-    },
-    NA
+    success <- callr::r(function() {
+      options(CITES_DB_DIR = "localdb")
+      con <- DBI::dbConnect(
+        MonetDBLite::MonetDBLite(),
+        getOption("CITES_DB_DIR")
+      )
+      out <- inherits(con, "MonetDBEmbeddedConnection")
+      citesdb::cites_disconnect()
+      options(CITES_DB_DIR = NULL)
+      return(out)
+    })
+    stopifnot(success)
+  },
+  NA
   )
 })
 

### 2019-03-18 README.Rmd
Add citation information

@@
   purrr::map_chr(~paste(.$given, .$family)) %>%
-  glue::glue_collapse(sep = ", ", last = ", and ") %>%
+  glue::glue_collapse(sep = ", ", last = " and ") %>%
   paste0("Authors: _", ., "_") %>%
@@
 
+### Citation
+
+If you use **citesdb** in a publication, please cite both the package and source data:
+
+```{r, results = "asis"}
+print(citation("citesdb"), style = "textVersion")
+```
+
 ### Contributing

### 2019-03-18 README.Rmd
Hide citation code in README
Add reference to rcites/Speciesplus

@@
 
+### Related work
+
+The [**rcites**](https://github.com/ropensci/rcites) package provides
+access to the Speciesplus/CITES Checklist API, which includes metadata about
+species and their protected status through time.
+
 ### Citation
@@
 
-```{r, results = "asis"}
+```{r, results = "asis", echo = FALSE}
 print(citation("citesdb"), style = "textVersion")

### 2019-03-18 README.Rmd
Download data if needed in README/vignette

@@
 options(width=120)
+knitr::opts_chunk$set(cache = TRUE)
 ```
@@
 results with `collect()`. For example:
+
+```{r getdata, include = FALSE}
+if (!citesdb::cites_status()) citesdb::cites_db_download()
+```
 

### 2019-03-18 README.Rmd
Add pkgdown

@@
 knitr::opts_chunk$set(cache = TRUE)
+citesdb::cites_disconnect()
+MonetDBLite::monetdblite_shutdown()
 ```
@@
 
+```{r message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
+cites_disconnect()
+```
+
 ### Related work

### 2019-03-18 tests/testthat/test-04-data.R
Use tmpfile for testing data download

@@ -2,9 +2,9 @@ context("Data")
 
 test_that("CITES data source hasn't changed", {
   skip_on_cran()
+  tmpf <- tempfile()
   httr::GET("https://trade.cites.org/cites_trade/download_db",
-            httr::write_disk("cites_trade_db.zip"))
-  expect_equivalent(tools::md5sum("cites_trade_db.zip"),
-                    "3d774b109c3ebf3594cf0fd4c20c1d1b")
-  unlink("cites_trade_db.zip")
+            httr::write_disk(tmpf))
+  expect_equivalent(tools::md5sum(tmpf), "3d774b109c3ebf3594cf0fd4c20c1d1b")
+  unlink(tmpf)
 })

### 2019-03-18 R/cites-guidance.R
removing files from gh-pages

@@ -1,83 +1,0 @@
-#' @title  CITES Trade Database shipment-level guidance
-#'
-#' @description _This text copied from the supplementary information file
-#'   accompanying the full database download from <https://trade.cites.org/>_
-#'
-#' @section Recommended citation:
-#'
-#'   UNEP-WCMC (Comps.) 2019. Full CITES Trade Database Download. Version
-#'   2019.2. CITES Secretariat, Geneva, Switzerland. Compiled by UNEP-WCMC,
-#'   Cambridge, UK. Available at: trade.cites.org.
-#'
-#' @section Background:
-#'
-#'   Under Article VIII of the Convention, Parties are required to provide
-#'   information regarding their trade in CITES-listed specimens through their
-#'   annual reports, and the Secretariat makes this information available
-#'   through the CITES Trade Database (trade.cites.org). This database currently
-#'   contains over 20 million records of international trade in CITES-listed
-#'   species. Parties recognise the importance of these reports as a tool for
-#'   monitoring the implementation of the Convention, assessing the
-#'   effectiveness of their wildlife management and trade policies, and to
-#'   enhance the detection of potentially harmful or illicit trade.
-#'
-#'   At the 70th meeting of the Standing Committee, Parties agreed that a full
-#'   non-aggregated version of the CITES Trade Database should be made available
-#'   and updated twice a year. These files represent the periodic release of the
-#'   CITES trade data in a shipment-by-shipment format with unique identifiers
-#'   replacing confidential permit numbers (see [SC70 Doc
-#'   26.2](https://cites.org/sites/default/files/eng/com/sc/70/E-SC70-26-02.pdf
-#'   ) and [SC70 Inf.
-#'   1](https://cites.org/sites/default/files/eng/com/sc/70/Inf/E-SC70-Inf-01.pdf)
-#'    for further background).
-#'
-#' @section Overview of data:
-#'
-#'   This zip file contains all trade records (including all historic data)
-#'   entered in the CITES Trade Database by 29 January 2019 and extracted at the
-#'   shipment level on 30 January 2019. This file is 2019.v2 and replaces
-#'   version 2019.v1 which contained some formatting anomalies and strengthens
-#'   security.
-#'
-#'   While the data provided through the search function on the [Web Portal of
-#'   the CITES Trade Database](https://trade.cites.org/) are aggregated, the
-#'   database contains non-aggregated data. The data provided in this download
-#'   is on a per-shipment basis i.e. it provides the relevant information about
[…]

### 2019-03-18 R/citesdb-package.R

@@ -1,25 +1,0 @@
-#' The CITES database package
-#'
-#' **citesdb** provides over 40 years of trade data from the [Convention on
-#' International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
-#' These data on imports and exports of CITES annex species are provided by
-#' CITES member countries to the CITES secretariat.
-#'
-#' This package allows you to download the full database of historic transactions
-#' to analyze locally on your computer. Because of the large size of this data,
-#' the **citesdb** package stores it as an on-disk [MonetDB](https://www.monetdb.org/)
-#' that can be queried without loading into RAM.
-#'
-#' @author Noam Ross, Evan Eskew, and Carlos Zambrana-Torrelio, [EcoHealth Alliance](https://www.ecohealthalliance.org/)
-#'
-#'
-#
-
-#' @keywords internal
-"_PACKAGE"
-
-# The following block is used by usethis to automatically manage
-# roxygen namespace tags. Modify with care!
-## usethis namespace: start
-## usethis namespace: end
-NULL

### 2019-03-18 R/connect.R

@@ -1,154 +1,0 @@
-#' @importFrom rappdirs user_data_dir
-cites_path <- function() {
-  sys_cites_path <- Sys.getenv("CITES_DB_DIR")
-  if (sys_cites_path == "") {
-    return(rappdirs::user_data_dir("citesdb"))
-  } else {
-    return(sys_cites_path)
-  }
-}
-
-check_status <- function() {
-  if (!cites_status(FALSE)) {
-    stop("Local CITES database empty or corrupt. Download with cites_db_download()") # nolint
-  }
-}
-
-#' The local CITES database
-#'
-#' Returns a connection to the local CITES database.  This is a DBI-compliant
-#' [MonetDBLite::MonetDBLite()] database connection.
-#'
-#' @param dbdir The location of the database on disk.  Defaults to
-#' `citesdb` under [rappdirs::user_data_dir()], or the environment variable `CITES_DB_DIR`.
-#'
-#' @return A MonetDBLite DBI connection
-#' @importFrom DBI dbIsValid dbConnect
-#' @importFrom MonetDBLite MonetDBLite
-#' @export
-#'
-#' @examples
-#' cites_db()
-cites_db <- function(dbdir = cites_path()) {
-  db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
-  if (inherits(db, "DBIConnection")) {
-    if (DBI::dbIsValid(db)) {
-      return(db)
-    }
-  }
-  dbname <- dbdir
-  dir.create(dbname, FALSE)
-
-  tryCatch(
-    db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir),
-    error = function(e) {
-      if (grepl("Database lock", e)) {
-        stop(paste(
-          "Local taxadb database is locked by another R session.\n",
-          "Try closing or running cites_disconect() in that session."
-        ),
-        call. = FALSE
-        )
-      } else {
-        stop(e)
-      }
-    },
-    finally = NULL
-  )
-
-  assign("cites_db", db, envir = cites_cache)
-  db
-}
-
-
-#' CITES Shipment Data
-#'
-#' Returns a remote table with all CITES shipment data. Requires the dplyr and dbplyr packages.
-#' @return A dplyr remote tibble ([dplyr::tbl()])
-#' @export
-#'
-#' @examples
-#' if (cites_status()) {
-#'   cites_shipments()
-#' }
-#' @importFrom dplyr tbl
-cites_shipments <- function() {
-  check_status()
[…]

### 2019-03-18 R/connection-pane.R

@@ -1,87 +1,0 @@
-sql_action <- function() {
-  if (requireNamespace("rstudioapi", quietly = TRUE) &&
-    exists("documentNew", asNamespace("rstudioapi"))) {
-    contents <- paste(
-      "-- !preview conn=citesdb::cites_db()",
-      "",
-      "SELECT * FROM cites_shipments LIMIT 100",
-      "",
-      sep = "\n"
-    )
-
-    rstudioapi::documentNew(
-      text = contents, type = "sql",
-      position = rstudioapi::document_position(2, 40),
-      execute = FALSE
-    )
-  }
-}
-
-#' Open CITES database connection pane in RStudio
-#'
-#' This function launches the RStudio "Connection" pane to interactively
-#' explore the database.
-#'
-#' @return NULL
-#' @export
-#'
-#' @examples
-#' if (!is.null(getOption("connectionObserver")))
-#'   cites_pane()
-cites_pane <- function() {
-  observer <- getOption("connectionObserver")
-  if (!is.null(observer)) {
-    observer$connectionOpened(
-      type = "CITESDB",
-      host = "citesdb",
-      displayName = "CITES Transaction Tables",
-      icon = system.file("img", "eha_logo.png", package = "citesdb"),
-      connectCode = "citesdb::cites_pane()",
-      disconnect = citesdb::cites_disconnect,
-      listObjectTypes = function() {
-        list(
-          table = list(contains = "data")
-        )
-      },
-      listObjects = function(type = "datasets") {
-        tbls <- DBI::dbListTables(cites_db())
-        data.frame(
-          name = tbls,
-          type = rep("table", length(tbls)),
-          stringsAsFactors = FALSE
-        )
-      },
-      listColumns = function(table) {
-        res <- DBI::dbSendQuery(cites_db(),
-                                paste("SELECT * FROM", table, "LIMIT 1"))
-        on.exit(DBI::dbClearResult(res))
-        data.frame(
-          name = res@env$info$names, type = res@env$info$types,
-          stringsAsFactors = FALSE
-        )
-      },
-      previewObject = function(rowLimit, table) {  #nolint
-        DBI::dbGetQuery(cites_db(),
-                        paste("SELECT * FROM", table, "LIMIT", rowLimit))
-      },
-      actions = list(
-        Status = list(
-          icon = system.file("img", "cites-logo.png", package = "citesdb"),
-          callback = cites_status
-        ),
-        SQL = list(
[…]

### 2019-03-18 R/download.R

@@ -1,146 +1,0 @@
-#' Download the CITES database to your local computer
-#'
-#' This command downloads the CITES shipments database and populates a local
-#' database.  The download is large (~158 MB), and the database will be over
-#' 1 GB on disk.  During import over 3.5 GB of disk space may be used temporarily.
-#'
-#' The database is stored by default under [rappdirs::user_data_dir()], or its
-#' location can be set with the environment variable `CITES_DB_DIR`.
-#'
-#' @param tag What release tag of data to download. Defaults to the most recent.
-#' @param destdir Where to download the compressed file.
-#' @param cleanup Whether to delete the compressed file after loading into the database.
-#' @param verbose Whether to display messages and download progress
-#'
-#' @return NULL
-#' @export
-#' @importFrom DBI dbRemoveTable dbExistsTable dbCreateTable dbExecute dbWriteTable
-#' @importFrom R.utils gunzip
-#'
-#' @examples
-#' \donttest{
-#' \dontrun{
-#' cites_db_download()
-#' }
-#' }
-cites_db_download <- function(tag = NULL, destdir = tempdir(),
-                              cleanup = TRUE, verbose = interactive()) {
-  if (verbose) message("Downloading data...\n")
-  zfile <- get_gh_release_file("ecohealthalliance/citesdb",
-    tag_name = tag,
-    destdir = destdir, verbose = verbose
-  )
-  ver <- attr(zfile, "ver")
-  if (verbose) message("Decompressing and building local database...\n")
-  temp_tsv <- tempfile(fileext = ".tsv")
-  gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
-
-  tblname <- "cites_shipments"
-  if (dbExistsTable(cites_db(), tblname)) {
-    dbRemoveTable(cites_db(), tblname)
-  }
-
-
-  dbCreateTable(cites_db(), tblname, fields = cites_field_types)
-
-  suppressMessages(
-    dbExecute(
-      cites_db(),
-      paste0(
-        "COPY OFFSET 2 INTO ", tblname, " FROM '",
-        temp_tsv,
-        "' USING DELIMITERS '\t','\n','\"' NULL as 'NA'"
-      )
-    )
-  )
-
-  dbWriteTable(cites_db(), "cites_status", make_status_table(version = ver),
-    overwrite = TRUE
-  )
-
-  load_citesdb_metadata()
-
[…]

### 2019-03-18 data-raw/process_trade_db.R

@@ -1,71 +1,0 @@
-library(httr)
-library(here)
-library(readr)
-library(fs)
-library(purrr)
-library(dplyr)
-GET(
-  "https://trade.cites.org/cites_trade/download_db",
-  write_disk(here("data-raw", "Trade_database.zip"), overwrite = TRUE)
-)
-unzip(here("data-raw", "Trade_database.zip"),
-  overwrite = TRUE,
-  exdir = here("data-raw"),
-)
-
-tdb <- map_dfr(
-  dir_ls(here("data-raw"), regexp = "^trade_db_\\d+\\.csv$"),
-  ~ read_csv(.,
-    col_types = cols(
-      Year = col_integer(),
-      Appendix = col_character(),
-      Taxon = col_character(),
-      Class = col_character(),
-      Order = col_character(),
-      Family = col_character(),
-      Genus = col_character(),
-      Term = col_character(),
-      Quantity = col_double(),
-      Unit = col_character(),
-      Importer = col_character(),
-      Exporter = col_character(),
-      Origin = col_character(),
-      Purpose = col_character(),
-      Source = col_character(),
-      Reporter.type = col_character(),
-      Import.permit.RandomID = col_character(),
-      Export.permit.RandomID = col_character(),
-      Origin.permit.RandomID = col_character()
-    )
-  )
-)
-
-# Removing invalid records to be fixed if possible in the future
-tdb <- tdb %>%
-  filter(!is.na(Year) & Year > 1970 & Year < 2020)
-
-# Arrange values; improves compressibility
-tdb <- tdb %>%
-  arrange(Year, Taxon, Order, Family, Genus, Term, Importer, Exporter, Appendix)
-
-# Write compressed data
-write_tsv(
-  tdb,
-  here("data-raw", "cites_trade_db.tsv.bz2")
-)
-
-# Clean up
-file_delete(
-  dir_ls(here("data-raw"), regexp = "(zip|csv|docx)$")
-)
-# Release the compressed data
-
-datastorr::github_release_info(
-  "ecohealthalliance/citesdb",
-  read = read_tsv,
-  filename = "cites_trade_db.tsv.bz2"
-) %>%
-datastorr::github_release_create(
-  description = "Release of CITES shipment data (v2019.2)",
-  target = "master", ignore_dirty = TRUE
-)

### 2019-03-18 tests/testthat/test-01-download.R

@@ -1,14 +1,0 @@
-context("Download")
-
-olddir <- Sys.getenv("CITES_DB_DIR")
-Sys.setenv(CITES_DB_DIR = normalizePath(file.path(getwd(), "localdb"),
-  mustWork = FALSE
-))
-
-test_that("Download succeeds", {
-  skip_on_cran()
-  cites_db_download()
-  expect_true(cites_status())
-})
-
-Sys.setenv(CITES_DB_DIR = olddir)

### 2019-03-18 tests/testthat/test-02-tables.R

@@ -1,41 +1,0 @@
-olddir <- Sys.getenv("CITES_DB_DIR")
-Sys.setenv(CITES_DB_DIR = normalizePath(file.path(getwd(), "localdb"),
-  mustWork = FALSE
-))
-
-context("Tables")
-
-test_that("Tables have expected types", {
-  skip_on_cran()
-  skip_if_not(cites_status())
-  expect_is(cites_shipments(), "tbl_MonetDBEmbeddedConnection")
-  expect_is(cites_metadata(), "data.frame")
-  expect_is(cites_codes(), "data.frame")
-  expect_is(cites_parties(), "data.frame")
-})
-
-test_that("All codes are accounted for", {
-  skip_on_cran()
-  skip_if_not(cites_status())
-  suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
-
-  sources <- cites_shipments() %>%
-    count(Source) %>%
-    pull(Source) %>%
-    na.omit()
-  source_codes <- cites_codes() %>%
-    filter(field == "Source") %>%
-    pull(code)
-  expect_setequal(sources, source_codes)
-
-  purposes <- cites_shipments() %>%
-    count(Purpose) %>%
-    pull(Purpose) %>%
-    na.omit()
-  purpose_codes <- cites_codes() %>%
-    filter(field == "Purpose") %>%
-    pull(code)
-  expect_setequal(purposes, purpose_codes)
-})
-
-Sys.setenv(CITES_DB_DIR = olddir)

### 2019-03-18 tests/testthat/test-03-connection.R

@@ -1,64 +1,0 @@
-olddir <- Sys.getenv("CITES_DB_DIR")
-Sys.setenv(CITES_DB_DIR = normalizePath(file.path(getwd(), "localdb"),
-                                        mustWork = FALSE
-))
-
-context("Connection")
-
-test_that("Database lock gives informative error", {
-  skip_on_cran()
-  skip_if_not(cites_status())
-  cites_db()
-  expect_error({
-    callr::r(function() {
-      options(CITES_DB_DIR = "localdb")
-      citesdb::cites_db()
-    })
-  },
-  "Local taxadb database is locked by another R session"
-  )
-})
-
-test_that("Disconnetion works", {
-  skip_on_cran()
-  skip_if_not(cites_status())
-
-  cites_disconnect()
-  expect_error({
-    success <- callr::r(function() {
-      options(CITES_DB_DIR = "localdb")
-      con <- DBI::dbConnect(
-        MonetDBLite::MonetDBLite(),
-        getOption("CITES_DB_DIR")
-      )
-      out <- inherits(con, "MonetDBEmbeddedConnection")
-      citesdb::cites_disconnect()
-      options(CITES_DB_DIR = NULL)
-      return(out)
-    })
-    stopifnot(success)
-  },
-  NA
-  )
-})
-
-test_that("Database is deleted", {
-  skip_on_cran()
-  skip_if_not(cites_status())
-
-  expect_error(cites_db_delete(), NA)
-  expect_equal(DBI::dbListTables(cites_db()), character(0))
-  expect_false(cites_status())
-})
-
-test_that("Tables fail when database is deleted", {
-  skip_on_cran()
-
-  expect_error(cites_shipments())
-  expect_error(cites_codes())
-  expect_error(cites_metadata())
-  expect_error(cites_parties())
-})
-
-unlink(Sys.getenv("CITES_DB_DIR"), recursive = TRUE)
-Sys.setenv(CITES_DB_DIR = olddir)

### 2019-03-18 tests/testthat/test-04-data.R

@@ -1,10 +1,0 @@
-context("Data")
-
-test_that("CITES data source hasn't changed", {
-  skip_on_cran()
-  tmpf <- tempfile()
-  httr::GET("https://trade.cites.org/cites_trade/download_db",
-            httr::write_disk(tmpf))
-  expect_equivalent(tools::md5sum(tmpf), "3d774b109c3ebf3594cf0fd4c20c1d1b")
-  unlink(tmpf)
-})

### 2019-03-18 README.Rmd
Change README/vignette header levels, fix deploy
Deploy tweaks for images

@@
 cites_disconnect()
+MonetDBLite::monetdblite_shutdown()
 ```
@@
 
-### Citation
+## Citation
 
@@
 
-### Contributing
+## Contributing
 
@@
 
-[![Created by EcoHealth Alliance](https://github.com/ecohealthalliance/citesdb/blob/master/inst/img/eha-footer.png)](https://www.ecohealthalliance.org/)
+[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ecohealthalliance/citesdb/master/man/figures/eha-footer.png)](https://www.ecohealthalliance.org/)

### 2019-03-18 README.Rmd
Fix yaml

@@
 
-[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ecohealthalliance/citesdb/master/man/figures/eha-footer.png)](https://www.ecohealthalliance.org/)
+[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ecohealthalliance/citesdb/master/man/figures/eha-footer.png){data-external="1"}](https://www.ecohealthalliance.org/)

### 2019-03-19 README.Rmd
Header change in README
Remove "on ordinary laptop" from README
As README is built on CI sometimes, this is no longer accurate.
Add note about loading whole DB into memory
Minor README updates, cites_parties() tweaks

@@
 ```{r setup, include = FALSE}
+
 knitr::opts_chunk$set(
@@
 )
+
 library(dplyr)
 ```
+
 # citesdb
@@
 ```{r authors, echo = FALSE, results = 'asis'}
+
 unclass(desc::desc_get_authors(here::here("DESCRIPTION"))) %>%
@@
 ```{r install_me, eval = FALSE}
+
 source("https://install-github.me/ecohealthalliance/citesdb")
@@
 
-```{r message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
+```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
+
 options(width=120)
@@
 ```{r getdata, include = FALSE}
+
 if (!citesdb::cites_status()) citesdb::cites_db_download()
@@
 ```{r, warning = FALSE}
+
 library(citesdb)
 library(dplyr)
+
 start <- Sys.time()
@@
 
+(_Note that running `collect()` on all of `cites_shipments()` will load a >3GB data frame into memory!_)
+
 The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making such
 analyses quite snappy even on such large data using normal desktops and laptops. Here's the
-timing of the above query, which processes over 20 million records, on an ordinary laptop:
+timing of the above query, which processes over 20 million records:
 
 ```{r}
+
 stop - start
@@
 ```{r}
+
 head(cites_metadata())
@@
 head(cites_codes())
+
+head(cites_parties())
 ```
@@
 
-```{r message=FALSE, warning=FALSE, error=FALSE, include=FALSE}
+### Basic analyses
+
+```{r warning = FALSE}
+
+library(ggplot2)
+library(cowplot)
+
+breaks <- seq(from = 1975, to = 2020, by = 5)
+
+plot1 <- cites_shipments() %>%
+  group_by(Year) %>%
+  summarize(n_records = n()) %>%
+  mutate(log10_n_records = log10(n_records)) %>%
+  ggplot(aes(x = Year, y = n_records)) +
+  geom_point() +
+  geom_line(linetype = "dashed", size = 0.2) +
+  ylab("Number of Records") +
+  theme_minimal() +
+  scale_x_continuous(breaks = breaks)
+
+plot2 <- cites_shipments() %>%
+  group_by(Year) %>%
+  summarize(n_records = n()) %>%
+  mutate(log10_n_records = log10(n_records)) %>%
+  ggplot(aes(x = Year, y = log10_n_records)) +
+  geom_point() +
+  geom_line(linetype = "dashed", size = 0.2) +
+  ylab("log10(Number of Records)") +
+  ylim(1.5, 6.5) +
+  theme_minimal() +
+  scale_x_continuous(breaks = breaks)
+
+plot_grid(plot1, plot2, nrow = 2)
+```
+
+```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
+
 cites_disconnect()
@@
 
-### Related work
+## Related work
 
@@
 ```{r, results = "asis", echo = FALSE}
+
 print(citation("citesdb"), style = "textVersion")

### 2019-03-19 R/connect.R

@@ -95,7 +95,10 @@ cites_shipments <- function() {
 #' @export
 #'
 #' @importFrom DBI dbReadTable
+#' @importFrom dplyr arrange
 #' @importFrom dplyr as_tibble
+#' @importFrom tidyr gather
+#' @importFrom tidyr separate
 #' @aliases metadata cites_metadata
 #' @examples
 #' if (cites_status()) {
@@ -118,14 +121,29 @@ cites_metadata <- function() {
 #' @rdname cites_metadata
 cites_codes <- function() {
   check_status()
-  as_tibble(dbReadTable(cites_db(), "cites_codes"))
+  arrange(as_tibble(dbReadTable(cites_db(), "cites_codes")), field, code)
 }
 
 #' @export
 #' @rdname cites_metadata
 cites_parties <- function() {
   check_status()
-  as_tibble(dbReadTable(cites_db(), "cites_metadata"))
+  raw_table <- dbReadTable(cites_db(), "cites_parties")
+  mod_table <- raw_table
+  mod_table[mod_table$country == "Namibia", "code"] <- "NA"
+  mod_table <- separate(mod_table, code,
+                        into = c("code1", "code2"),
+                        sep = ", formerly| ex-",
+                        fill = "right")
+  mod_table$code1 <- gsub(" *", "", mod_table$code1)
+  mod_table$code2 <- gsub(" *", "", mod_table$code2)
+  mod_table <- gather(mod_table, key = code_version, value = code, -country, -date)
+  mod_table <- mod_table[!is.na(mod_table$code), ]
+  mod_table$code_version <- ifelse(mod_table$code_version == "code2", TRUE, FALSE)
+  mod_table <- mod_table[ , c(1, 4, 3, 2)]
+  colnames(mod_table) <- c("country", "code", "former_code", "date")
+  mod_table <- arrange(mod_table, country, desc(former_code))
+  as_tibble(mod_table)
 }
 
 #' Disconnect from the CITES database

### 2019-03-19 R/connect.R
Moving table processing to data-raw

@@ -95,10 +95,7 @@ cites_shipments <- function() {
 #' @export
 #'
 #' @importFrom DBI dbReadTable
-#' @importFrom dplyr arrange
 #' @importFrom dplyr as_tibble
-#' @importFrom tidyr gather
-#' @importFrom tidyr separate
 #' @aliases metadata cites_metadata
 #' @examples
 #' if (cites_status()) {
@@ -121,29 +118,14 @@ cites_metadata <- function() {
 #' @rdname cites_metadata
 cites_codes <- function() {
   check_status()
-  arrange(as_tibble(dbReadTable(cites_db(), "cites_codes")), field, code)
+  as_tibble(dbReadTable(cites_db(), "cites_codes"))
 }
 
 #' @export
 #' @rdname cites_metadata
 cites_parties <- function() {
   check_status()
-  raw_table <- dbReadTable(cites_db(), "cites_parties")
-  mod_table <- raw_table
-  mod_table[mod_table$country == "Namibia", "code"] <- "NA"
-  mod_table <- separate(mod_table, code,
-                        into = c("code1", "code2"),
-                        sep = ", formerly| ex-",
-                        fill = "right")
-  mod_table$code1 <- gsub(" *", "", mod_table$code1)
-  mod_table$code2 <- gsub(" *", "", mod_table$code2)
-  mod_table <- gather(mod_table, key = code_version, value = code, -country, -date)
-  mod_table <- mod_table[!is.na(mod_table$code), ]
-  mod_table$code_version <- ifelse(mod_table$code_version == "code2", TRUE, FALSE)
-  mod_table <- mod_table[ , c(1, 4, 3, 2)]
-  colnames(mod_table) <- c("country", "code", "former_code", "date")
-  mod_table <- arrange(mod_table, country, desc(former_code))
-  as_tibble(mod_table)
+  as_tibble(dbReadTable(cites_db(), "cites_parties"))
 }
 
 #' Disconnect from the CITES database

### 2019-03-19 data-raw/process_tables.R

@@ -1,0 +1,49 @@
+
+
+library(tidyverse)
+library(tabulizer)
+
+h <- here::here
+
+#==============================================================================
+
+
+# Prepare cites_metadata table
+
+cites_metadata <- read_tsv(h("data-raw", "cites_metadata.tsv"))
+
+write_tsv(cites_metadata, h("inst", "extdata", "cites_metadata.tsv"))
+
+#==============================================================================
+
+
+# Prepare cites_codes table
+
+cites_codes <- read_tsv(h("data-raw", "cites_codes.tsv")) %>%
+  arrange(field, code) %>%
+  filter(field != "Term", field != "Unit")
+
+# Note, currently leaving out "Term" and "Unit" since these variables appear
+# as full descriptions (not codes) in the citesdb data
+
+write_tsv(cites_codes, h("inst", "extdata", "cites_codes.tsv"))
+
+#==============================================================================
+
+
+# Prepare cites_parties table
+
+cites_parties <- read_tsv(h("data-raw", "cites_parties.tsv")) %>%
+  mutate(code = ifelse(country == "Namibia", "NA", code)) %>%
+  separate(., code, into = c("code1", "code2"),
+           sep = ", formerly| ex-",fill = "right") %>%
+  mutate(code1 = str_replace_all(code1, " ", ""),
+         code2 = str_replace_all(code2, " ", "")) %>%
+  gather(., key = code_version, value = code, -country, -date) %>%
+  filter(!is.na(code)) %>%
+  mutate(code_version = ifelse(code_version == "code2", TRUE, FALSE)) %>%
+  select(country, code, code_version, date) %>%
+  rename(former_code = code_version) %>%
+  arrange(country, desc(former_code))
+
+write_tsv(cites_parties, h("inst", "extdata", "cites_parties.tsv"))

### 2019-03-20 R/connect.R
text changes

@@ -44,7 +44,7 @@ cites_db <- function(dbdir = cites_path()) {
     error = function(e) {
       if (grepl("Database lock", e)) {
         stop(paste(
-          "Local taxadb database is locked by another R session.\n",
+          "Local citesdb database is locked by another R session.\n",
           "Try closing or running cites_disconect() in that session."
         ),
         call. = FALSE

### 2019-03-20 tests/testthat/test-03-connection.R

@@ -15,7 +15,7 @@ test_that("Database lock gives informative error", {
       citesdb::cites_db()
     })
   },
-  "Local taxadb database is locked by another R session"
+  "Local citesdbs database is locked by another R session"
   )
 })
 

### 2019-03-20 data-raw/process_tables.R
Completing country coverage in cites_parties()

@@ -10,7 +10,8 @@ h <- here::here
 
 # Prepare cites_metadata table
 
-cites_metadata <- read_tsv(h("data-raw", "cites_metadata.tsv"))
+cites_metadata <- h("data-raw", "cites_metadata.tsv") %>%
+  read_tsv()
 
 write_tsv(cites_metadata, h("inst", "extdata", "cites_metadata.tsv"))
 
@@ -19,7 +20,8 @@ write_tsv(cites_metadata, h("inst", "extdata", "cites_metadata.tsv"))
 
 # Prepare cites_codes table
 
-cites_codes <- read_tsv(h("data-raw", "cites_codes.tsv")) %>%
+cites_codes <- h("data-raw", "cites_codes.tsv") %>%
+  read_tsv() %>%
   arrange(field, code) %>%
   filter(field != "Term", field != "Unit")
 
@@ -33,7 +35,8 @@ write_tsv(cites_codes, h("inst", "extdata", "cites_codes.tsv"))
 
 # Prepare cites_parties table
 
-cites_parties <- read_tsv(h("data-raw", "cites_parties.tsv")) %>%
+cites_parties <- h("data-raw", "cites_parties.tsv") %>%
+  read_tsv() %>%
   mutate(code = ifelse(country == "Namibia", "NA", code)) %>%
   separate(., code, into = c("code1", "code2"),
            sep = ", formerly| ex-", fill = "right") %>%
@@ -42,21 +45,20 @@ cites_parties <- read_tsv(h("data-raw", "cites_parties.tsv")) %>%
   gather(., key = code_version, value = code, -country, -date) %>%
   filter(!is.na(code)) %>%
   mutate(code_version = ifelse(code_version == "code2", TRUE, FALSE)) %>%
-  select(country, code, code_version, date) %>%
   rename(former_code = code_version) %>%
+  select(country, code, former_code, date) %>%
+  mutate(data_source = "'A guide to using the CITES Trade Database', Version 8, Annex 4") %>%
   filter(!(country == "Slovakia" & former_code == TRUE),
          !(country == "Czech Republic" & former_code == TRUE))
 
-countries_raw <- h("data-raw", "en-CITES_Trade_Database_Guide.pdf") %>%
+countries_raw <- h("data-raw", "country_code_data", "en-CITES_Trade_Database_Guide.pdf") %>%
   extract_areas(., pages = c(14, 15, 16))
 
[…]

### 2019-03-20 tests/testthat/test-03-connection.R
Typo fix to address failing CI test

@@ -15,7 +15,7 @@ test_that("Database lock gives informative error", {
       citesdb::cites_db()
     })
   },
-  "Local citesdbs database is locked by another R session"
+  "Local citesdb database is locked by another R session"
   )
 })
 

### 2019-04-03 README.Rmd
Moving table processing to data-raw
text changes
Completing country coverage in cites_parties()
README updates, begin vignette revision

@@
 
-**citesdb** is a package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>.
-This data consists of over 20 million records of reported shipments of wildlife and wildlife
-products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
+**citesdb** is an R package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>.
+This data consists of over 20 million records of reported shipments of wildlife and wildlife products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
 
@@
 
-options(width=120)
+options(width = 120)
 knitr::opts_chunk$set(cache = TRUE)
@@
 
-Not to worry, just do as it says and run `cites_db_download()`.  This will
-fetch the most recent database from online, an approximately 158 MB download.
-It will expand to over 1 GB in the local database. During
-the download and database building up to 3.5 GB of disk space may be used temporarily.
+Not to worry, just do as it says and run `cites_db_download()`.
+This will fetch the most recent database from online, an approximately 158 MB download.
+It will expand to over 1 GB in the local database.
+During the download and database building up to 3.5 GB of disk space may be used temporarily.
 
@@
 
-Once you fetch the data you can connect to the database with the `cites_db()`
-command.  You can use the
-`cites_shipments()` command to load a remote `tibble` that is backed by the
-database but not loaded into R.  You can use this
-to analyze CITES data without ever loading it into memory, then gather your
-results with `collect()`. For example:
+Once you fetch the data you can connect to the database with the `cites_db()` command.
+You can use the `cites_shipments()` command to load a remote `tibble` that is backed by the database but not loaded into R.
+You can use this to analyze CITES data without ever loading it into memory, then gather your results with the `dplyr` function `collect()`.
+For example:
 
@@
   group_by(Year) %>%
-  summarize(number = n()) %>%
+  summarize(n_records = n()) %>%
   arrange(desc(Year)) %>%
@@
 
-(_Note that running `collect()` on all of `cites_shipments()` will load a >3GB data frame into memory!_)
+(_Note that running `collect()` on all of `cites_shipments()` will load a >3 GB data frame into memory!_)
 
-The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making such
-analyses quite snappy even on such large data using normal desktops and laptops. Here's the
-timing of the above query, which processes over 20 million records:
+The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making such analyses quite snappy even on such large data using normal desktops and laptops.
+Here's the timing of the above query, which processes over 20 million records:
 
@@
 
-If you are using a recent version of RStudio interactively, loading the CITES package also brings
-up a browsable pane in the "Connections" tab that lets you explore and preview
-the database, as well as interact with it directly via SQL commands.
+If you are using a recent version of RStudio interactively, loading the CITES package also brings up a browsable pane in the "Connections" tab that lets you explore and preview the database, as well as interact with it directly via SQL commands.
 
-If you don't need any of the bells and whistles of this package, you can download
-the raw data as a single compressed TSV from the [releases page](https://github.com/ecohealthalliance/citesdb/releases),
-or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.
+If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.
 
@@
 
-The database also contains tables of field metadata, codes used, and CITES countries. This information
-comes from ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf),
-on the CITES website.  Convenience functions `cites_metadata()`, `cites_codes()`, and `cites_parties()`
-access this information:
+The package database also contains tables of field metadata, codes used, and CITES countries.
+This information comes from ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf), on the CITES website.
+Convenience functions `cites_metadata()`, `cites_codes()`, and `cites_parties()` access this information:
 
@@
 
-More information on the release of shipment-level data can be found in the
-`?guidance` help file.
-
-### Basic analyses
-
-```{r warning = FALSE}
-
-library(ggplot2)
-library(cowplot)
-
-breaks <- seq(from = 1975, to = 2020, by = 5)
-
-plot1 <- cites_shipments() %>%
-  group_by(Year) %>%
-  summarize(n_records = n()) %>%
-  mutate(log10_n_records = log10(n_records)) %>%
-  ggplot(aes(x = Year, y = n_records)) +
-  geom_point() +
-  geom_line(linetype = "dashed", size = 0.2) +
-  ylab("Number of Records") +
-  theme_minimal() +
-  scale_x_continuous(breaks = breaks)
-
-plot2 <- cites_shipments() %>%
-  group_by(Year) %>%
-  summarize(n_records = n()) %>%
-  mutate(log10_n_records = log10(n_records)) %>%
-  ggplot(aes(x = Year, y = log10_n_records)) +
-  geom_point() +
-  geom_line(linetype = "dashed", size = 0.2) +
-  ylab("log10(Number of Records)") +
-  ylim(1.5, 6.5) +
-  theme_minimal() +
-  scale_x_continuous(breaks = breaks)
-
-plot_grid(plot1, plot2, nrow = 2)
-```
-
 ```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
 
-cites_disconnect()
+citesdb::cites_disconnect()
 MonetDBLite::monetdblite_shutdown()
@@
 
+More information on the release of shipment-level CITES data can be found in the `?guidance` help file.
+
 ## Related work
 
-The [**rcites**](https://github.com/ropensci/rcites) package provides
-access to the Speciesplus/CITES Checklist API, which includes metadata about
-species and their protected status through time.
+The [**rcites**](https://github.com/ropensci/rcites) package provides access to the Speciesplus/CITES Checklist API, which includes metadata about species and their protected status through time.
 
@@
 
-Want have feedback or want to contribute?
+Have feedback or want to contribute?
 Great!

### 2019-04-05 R/connect.R
Minor typo and formatting fixes

@@ -16,10 +16,10 @@ check_status <- function() {
 
 #' The local CITES database
 #'
-#' Returns a connection to the local CITES database.  This is a DBI-compliant
+#' Returns a connection to the local CITES database. This is a DBI-compliant
 #' [MonetDBLite::MonetDBLite()] database connection.
 #'
-#' @param dbdir The location of the database on disk.  Defaults to
+#' @param dbdir The location of the database on disk. Defaults to
 #' `citesdb` under [rappdirs::user_data_dir()], or the environment variable `CITES_DB_DIR`.
 #'
 #' @return A MonetDBLite DBI connection
@@ -61,7 +61,7 @@ cites_db <- function(dbdir = cites_path()) {
 }
 
 
-#' CITES Shipment Data
+#' CITES shipment data
 #'
 #' Returns a remote table with all CITES shipment data. Requires the dplyr and dbplyr packages.
 #' @return A dplyr remote tibble ([dplyr::tbl()])
@@ -77,18 +77,18 @@ cites_shipments <- function() {
   tbl(cites_db(), "cites_shipments")
 }
 
-#' Cites shipment metadata
+#' CITES shipment metadata
 #'
 #' @description
 #'
 #' The CITES database also includes tables of column-level metadata and
-#' meanings of codes in columns, as well as a listing of CITES parties.
-#' convenience functions access these tables. As they are small, the functions
-#' collect the into R session memory, rather than returning a remote table.
+#' meanings of codes in columns, as well as a listing of CITES Parties/country abbreviations.
+#' Convenience functions access these tables. As they are small, the functions
+#' collect the tables into R session memory, rather than returning a remote table.
 #'
 #' This information is drawn from
 #' ["A guide to using the CITES Trade Database"](https://trade.cites.org/cites_trade_guidelines/en-CITES_Trade_Database_Guide.pdf),
-#' from the CITES website. More information the the shipment-level data can be
+#' from the CITES website. More information on the shipment-level data can be
 #' found in the [guidance] help file.
 #'
 #' @return A tibble of metadata

### 2019-04-05 R/download.R
Put link to releases in docs

@@ -8,6 +8,8 @@
 #' location can be set with the environment variable `CITES_DB_DIR`.
 #'
 #' @param tag What release tag of data to download. Defaults to the most recent.
+#'   Releases are expected to come twice per year.  See all releases at
+#'   <https://github.com/ecohealthalliance/citesdb/releases>.
 #' @param destdir Where to download the compressed file.
 #' @param cleanup Whether to delete the compressed file after loading into the database.
 #' @param verbose Whether to display messages and download progress

### 2019-04-05 README.Rmd
Update vignette

@@
 ```{r}
-
 head(cites_metadata())
@@
 
-[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ecohealthalliance/citesdb/master/man/figures/eha-footer.png){data-external="1"}](https://www.ecohealthalliance.org/)
+[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ecohealthalliance/citesdb/master/vignettes/figures/eha-footer.png){data-external="1"}](https://www.ecohealthalliance.org/)

### 2019-04-05 README.Rmd
Tweaks to various docs

@@
 ```{r setup, include = FALSE}
-
 knitr::opts_chunk$set(
@@
 ```{r authors, echo = FALSE, results = 'asis'}
-
 unclass(desc::desc_get_authors(here::here("DESCRIPTION"))) %>%
@@
 **citesdb** is an R package to conveniently analyze the full CITES shipment-level wildlife trade database, available at <https://trade.cites.org/>.
-This data consists of over 20 million records of reported shipments of wildlife and wildlife products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
+This data consists of over 40 years and 20 million records of reported shipments of wildlife and wildlife products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
 
@@
 ```{r install_me, eval = FALSE}
-
 source("https://install-github.me/ecohealthalliance/citesdb")
@@
 ```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
-
 options(width = 120)
@@
 ```{r getdata, include = FALSE}
-
 if (!citesdb::cites_status()) citesdb::cites_db_download()
@@
 ```{r, warning = FALSE}
-
 library(citesdb)
@@
 ```{r message = FALSE, warning = FALSE, error = FALSE, include = FALSE}
-
 citesdb::cites_disconnect()
@@
 ```{r, results = "asis", echo = FALSE}
-
 print(citation("citesdb"), style = "textVersion")

### 2019-04-05 R/citesdb-package.R

@@ -1,16 +1,17 @@
 #' The CITES database package
 #'
-#' **citesdb** provides over 40 years of trade data from the [Convention on
+#' **citesdb** provides over 40 years of wildlife trade data from the [Convention on
 #' International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
 #' These data on imports and exports of CITES annex species are provided by
 #' CITES member countries to the CITES secretariat.
 #'
 #' This package allows you to download the full database of historic transactions
+#' (commonly called the [CITES Trade Database](https://trade.cites.org/))
 #' to analyze locally on your computer. Because of the large size of this data,
 #' the **citesdb** package stores it as an on-disk [MonetDB](https://www.monetdb.org/)
 #' that can be queried without loading into RAM.
 #'
-#' @author Noam Ross, Evan Eskew, and Carlos Zambrana-Torrelio, [EcoHealth Alliance](https://www.ecohealthalliance.org/)
+#' @author Noam Ross and Evan A. Eskew, [EcoHealth Alliance](https://www.ecohealthalliance.org/)
 #'
 #'
 #

### 2019-04-05 R/download.R

@@ -1,15 +1,15 @@
 #' Download the CITES database to your local computer
 #'
 #' This command downloads the CITES shipments database and populates a local
-#' database.  The download is large (~158 MB), and the database will be over
-#' 1 GB on disk.  During import over 3.5 GB of disk space may be used temporarily.
+#' database. The download is large (~158 MB), and the database will be over
+#' 1 GB on disk. During import over 3.5 GB of disk space may be used temporarily.
 #'
 #' The database is stored by default under [rappdirs::user_data_dir()], or its
 #' location can be set with the environment variable `CITES_DB_DIR`.
 #'
 #' @param tag What release tag of data to download. Defaults to the most recent.
-#'   Releases are expected to come twice per year.  See all releases at
-#'   <https://github.com/ecohealthalliance/citesdb/releases>.
+#' Releases are expected to come twice per year. See all releases at
+#' <https://github.com/ecohealthalliance/citesdb/releases>.
 #' @param destdir Where to download the compressed file.
 #' @param cleanup Whether to delete the compressed file after loading into the database.
 #' @param verbose Whether to display messages and download progress

### 2019-04-05 data-raw/process_trade_db.R

@@ -1,13 +1,17 @@
+
+
 library(httr)
 library(here)
 library(readr)
 library(fs)
 library(purrr)
 library(dplyr)
+
 GET(
   "https://trade.cites.org/cites_trade/download_db",
   write_disk(here("data-raw", "Trade_database.zip"), overwrite = TRUE)
 )
+
 unzip(here("data-raw", "Trade_database.zip"),
   overwrite = TRUE,
   exdir = here("data-raw"),
@@ -58,8 +62,8 @@ write_tsv(
 file_delete(
   dir_ls(here("data-raw"), regexp = "(zip|csv|docx)$")
 )
-# Release the compressed data
 
+# Release the compressed data
 datastorr::github_release_info(
   "ecohealthalliance/citesdb",
   read = read_tsv,

### 2019-04-05 README.Rmd
Add Zenodo DOI and citation info

@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2630836.svg){data-external="1"}](https://doi.org/10.5281/zenodo.2630836)
 [![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=shield){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
@@
 ```{r}
-
 stop - start

### 2019-04-16 README.Rmd
Adding rOpenSci badge

--- 2019-04-05 README.Rmd
@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2630836.svg){data-external="1"}](https://doi.org/10.5281/zenodo.2630836)
 [![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=shield){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![](https://badges.ropensci.org/292_status.svg)](https://github.com/ropensci/software-review/issues/292)
 [![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)

### 2019-05-07 vignettes/app-data-process.R
remove duplicate reports; do not group by reporting country

@@ -6,7 +6,7 @@ library(sf)
 
 pangolin_shipments <- cites_shipments() %>%
   filter(Order == "Pholidota", Purpose %in% c(NA_character_, "T")) %>%
-  select(-Purpose, -Import.permit.RandomID, -Export.permit.RandomID, -Origin.permit.RandomID, -Class, -Order, -Family, -Genus) %>%
+  select(-Purpose, -Import.permit.RandomID, -Export.permit.RandomID, -Origin.permit.RandomID, -Class, -Order, -Family, -Genus, -Reporter.type ) %>%
   collect()
 cites_disconnect()
 
@@ -28,23 +28,24 @@ pangolin_shipments <- pangolin_shipments %>%
 
 dat <- pangolin_shipments %>%
   janitor::clean_names() %>%
-  mutate(reporting_country = ifelse(reporter_type == "I", importer, exporter),
-         reporting_country  = countrycode(sourcevar = reporting_country,
-                                          origin = "iso2c",
-                                          destination = "country.name"),
-         start2 = countrycode(sourcevar = start,
-                             origin = "iso2c",
-                             destination = "country.name"),
-         end = countrycode(sourcevar = importer,
-                           origin = "iso2c",
-                           destination = "country.name")) %>%
-  select(year, start2, end, reporting_country, term, start_lon, start_lat, end_lon, end_lat) %>%
+  mutate(
+    # reporting_country = ifelse(reporter_type == "I", importer, exporter),
+    # reporting_country  = countrycode(sourcevar = reporting_country,
+    #                                  origin = "iso2c",
+    #                                  destination = "country.name"),
+    start2 = countrycode(sourcevar = start,
+                         origin = "iso2c",
+                         destination = "country.name"),
+    end = countrycode(sourcevar = importer,
+                      origin = "iso2c",
+                      destination = "country.name")) %>%
+  select(year, start2, end, term, start_lon, start_lat, end_lon, end_lat) %>%
   rename(start = start2)
 
 
 dat2 <- dat %>%
   group_by_all() %>%
-  summarize(n = n()) %>%
+  mutate(n = n()) %>%
   ungroup()
 
 write_rds(dat2, here::here("vignettes", "pangolin_dat.rds"))

### 2019-05-07 R/connect.R
Install via devtools in README, check_status() to cites_check_status()

@@ -8,7 +8,7 @@ cites_path <- function() {
   }
 }
 
-check_status <- function() {
+cites_check_status <- function() {
   if (!cites_status(FALSE)) {
     stop("Local CITES database empty or corrupt. Download with cites_db_download()") # nolint
   }
@@ -73,7 +73,7 @@ cites_db <- function(dbdir = cites_path()) {
 #' }
 #' @importFrom dplyr tbl
 cites_shipments <- function() {
-  check_status()
+  cites_check_status()
   tbl(cites_db(), "cites_shipments")
 }
 
@@ -110,21 +110,21 @@ cites_shipments <- function() {
 #'   dplyr::tbl(cites_db(), "cites_parties")
 #' }
 cites_metadata <- function() {
-  check_status
+  cites_check_status()
   as_tibble(dbReadTable(cites_db(), "cites_metadata"))
 }
 
 #' @export
 #' @rdname cites_metadata
 cites_codes <- function() {
-  check_status()
+  cites_check_status()
   as_tibble(dbReadTable(cites_db(), "cites_codes"))
 }
 
 #' @export
 #' @rdname cites_metadata
 cites_parties <- function() {
-  check_status()
+  cites_check_status()
   as_tibble(dbReadTable(cites_db(), "cites_parties"))
 }
 

### 2019-05-07 R/download.R
cites_status() shows database location on disk

@@ -42,7 +42,6 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
     dbRemoveTable(cites_db(), tblname)
   }
 
-
   dbCreateTable(cites_db(), tblname, fields = cites_field_types)
 
   suppressMessages(
@@ -105,7 +104,8 @@ make_status_table <- function(version) {
       DBI::dbGetQuery(cites_db(),
                       "SELECT COUNT(*) FROM cites_shipments;")[[1]],
       format = "d", big.mark = ","),
-    size_on_disk = format(sz, "auto")
+    size_on_disk = format(sz, "auto"),
+    location_on_disk = cites_path()
   )
 }
 

### 2019-05-07 R/connection-pane.R
Formatting example in cites_pane()

@@ -26,8 +26,7 @@ sql_action <- function() {
 #' @export
 #'
 #' @examples
-#' if (!is.null(getOption("connectionObserver")))
-#'   cites_pane()
+#' if (!is.null(getOption("connectionObserver"))) cites_pane()
 cites_pane <- function() {
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {

### 2019-05-08 R/download.R
Disconnecct after deleting tables to trigger cleanup

@@ -16,7 +16,8 @@
 #'
 #' @return NULL
 #' @export
-#' @importFrom DBI dbRemoveTable dbExistsTable dbCreateTable dbExecute dbWriteTable
+#' @importFrom DBI dbRemoveTable dbExistsTable dbCreateTable dbExecute
+#'   dbWriteTable dbListTables
 #' @importFrom R.utils gunzip
 #'
 #' @examples
@@ -37,11 +38,12 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
   temp_tsv <- tempfile(fileext = ".tsv")
   gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
 
-  tblname <- "cites_shipments"
-  if (dbExistsTable(cites_db(), tblname)) {
-    dbRemoveTable(cites_db(), tblname)
+  for (tab in dbListTables(cites_db())) {
+    dbRemoveTable(cites_db(), tab)
   }
 
+  cites_disconnect()
+  tblname <- "cites_shipments"
   dbCreateTable(cites_db(), tblname, fields = cites_field_types)
 
   suppressMessages(

### 2019-05-08 tests/testthat/test-lint.R
Exclude linters that trigger false positives.
Closes #2

@@ -2,6 +2,11 @@ if (requireNamespace("lintr", quietly = TRUE)) {
   context("lints")
   test_that("Package Style", {
     skip_on_cran()
-    lintr::expect_lint_free()
+    lintr::expect_lint_free(
+      linters = lintr::with_defaults(
+        camel_case_linter = NULL,
+        snake_case_linter = NULL,
+        absolute_paths_linter = NULL
+      ))
   })
 }

### 2019-05-08 R/download.R
Move load_citesdb_metadata to download.R
More logical where it's called here.

@@ -148,3 +148,20 @@ get_gh_release_file <- function(repo, tag_name = NULL, destdir = tempdir(),
   attr(out_path, "ver") <- release_obj[[1]]$tag_name
   return(out_path)
 }
+
+#' @importFrom utils read.table
+load_citesdb_metadata <- function() {
+  tsvs <- list.files(system.file("extdata", package = "citesdb"),
+                     pattern = "\\.tsv$", full.names = TRUE
+  )
+  tblnames <- tools::file_path_sans_ext(basename(tsvs))
+  for (i in seq_along(tsvs)) {
+    suppressMessages(dbWriteTable(cites_db(), tblnames[i],
+                                  read.table(tsvs[i],
+                                             stringsAsFactors = FALSE, sep = "\t",
+                                             header = TRUE, quote = "\""
+                                  ),
+                                  overwrite = TRUE
+    ))
+  }
+}

### 2019-05-08 R/onattach.R

@@ -59,20 +59,3 @@ cites_status <- function(verbose = TRUE) {
   invisible(out)
 }
 
-
-#' @importFrom utils read.table
-load_citesdb_metadata <- function() {
-  tsvs <- list.files(system.file("extdata", package = "citesdb"),
-    pattern = "\\.tsv$", full.names = TRUE
-  )
-  tblnames <- tools::file_path_sans_ext(basename(tsvs))
-  for (i in seq_along(tsvs)) {
-    suppressMessages(dbWriteTable(cites_db(), tblnames[i],
-      read.table(tsvs[i],
-        stringsAsFactors = FALSE, sep = "\t",
-        header = TRUE, quote = "\""
-      ),
-      overwrite = TRUE
-    ))
-  }
-}

### 2019-05-08 R/connect.R
Simplify disconnection to avoid recursion

@@ -142,7 +142,7 @@ cites_disconnect <- function() {
 cites_disconnect_ <- function(environment = cites_cache) { # nolint
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
-    MonetDBLite::monetdblite_shutdown()
+    DBI::dbDisconnect(db)
   }
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {

### 2019-05-08 README.Rmd
Install via devtools in README, check_status() to cites_check_status()
Use remotes instead of devtools to install
Make CI badge reflect only master branch

--- 2019-04-05 README.Rmd
@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
-[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb.svg?style=shield){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2630836.svg){data-external="1"}](https://doi.org/10.5281/zenodo.2630836)
+[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb/tree/master.svg?style=shield){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![](https://badges.ropensci.org/292_status.svg)](https://github.com/ropensci/software-review/issues/292)
 [![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
@@
 ```{r install_me, eval = FALSE}
-source("https://install-github.me/ecohealthalliance/citesdb")
+remotes::install_github("ecohealthalliance/citesdb")
 ```

### 2019-05-08 inst/response_to_reviewers.md
Add response to reviewers to

# Response to reviewers

- We have changed the installation method to use `remotes` (which has lower
  dependencies than `devtools`).
- We have added disk location to `cites_status`
- We now clear out and disconnect the database tables after updating, which
  should trigger disk cleanup and avoid doubling database size

- We have opted not to use **glue** but stick with `paste` in the interest of
  limiting dependencies.  This is a trade-off but we believe a minor one.
- The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
  is due to tests skippeed on CRAN.  Setting the environment variable to 
  `NOT_CRAN=true` shows that our test coverage is 65%.  This is lower than
  typical, but as we note above, this is largely due to the verbose code
  for interacting with the RStudio connection pane, which can not be
  straightforwardly testeed. Other code coverage is 
  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
  and all major funcionality is tested.

### 2019-05-08 R/connect.R
shutdown when disconnecting from the db

@@ -17,7 +17,10 @@ cites_check_status <- function() {
 #' The local CITES database
 #'
 #' Returns a connection to the local CITES database. This is a DBI-compliant
-#' [MonetDBLite::MonetDBLite()] database connection.
+#' [MonetDBLite::MonetDBLite()] database connection. When using **dplyr**-based
+#' workflows, one typically acccesses tables with functions such as
+#' [cites_shipments()], but this function lets one interact with the database
+#' directly via SQL.
 #'
 #' @param dbdir The location of the database on disk. Defaults to
 #' `citesdb` under [rappdirs::user_data_dir()], or the environment variable `CITES_DB_DIR`.
@@ -28,7 +31,10 @@ cites_check_status <- function() {
 #' @export
 #'
 #' @examples
-#' cites_db()
+#' library(DBI)
+#' dbListTables(cites_db())
+#' parties <- dbReadTable(cites_db, "cites_parties")
+#'
 cites_db <- function(dbdir = cites_path()) {
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
@@ -142,7 +148,7 @@ cites_disconnect <- function() {
 cites_disconnect_ <- function(environment = cites_cache) { # nolint
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
-    DBI::dbDisconnect(db)
+    DBI::dbDisconnect(db, shutdown = TRUE)
   }
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {

### 2019-05-08 R/connect.R
Catch other versions of MoneeDB error

@@ -18,7 +18,7 @@ cites_check_status <- function() {
 #'
 #' Returns a connection to the local CITES database. This is a DBI-compliant
 #' [MonetDBLite::MonetDBLite()] database connection. When using **dplyr**-based
-#' workflows, one typically acccesses tables with functions such as
+#' workflows, one typically accesses tables with functions such as
 #' [cites_shipments()], but this function lets one interact with the database
 #' directly via SQL.
 #'
@@ -31,10 +31,13 @@ cites_check_status <- function() {
 #' @export
 #'
 #' @examples
+#' \donttest{
+#' \dontrun{
 #' library(DBI)
 #' dbListTables(cites_db())
 #' parties <- dbReadTable(cites_db, "cites_parties")
-#'
+#' }
+#' }
 cites_db <- function(dbdir = cites_path()) {
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
@@ -48,7 +51,7 @@ cites_db <- function(dbdir = cites_path()) {
   tryCatch(
     db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir),
     error = function(e) {
-      if (grepl("Database lock", e)) {
+      if (grepl("(Database lock|bad rolemask)", e)) {
         stop(paste(
           "Local citesdb database is locked by another R session.\n",
           "Try closing or running cites_disconect() in that session."

### 2019-05-08 R/connect.R
Fix example and documentation typo in cites_db

@@ -35,7 +35,7 @@ cites_check_status <- function() {
 #' \dontrun{
 #' library(DBI)
 #' dbListTables(cites_db())
-#' parties <- dbReadTable(cites_db, "cites_parties")
+#' parties <- dbReadTable(cites_db(), "cites_parties")
 #' }
 #' }
 cites_db <- function(dbdir = cites_path()) {

### 2019-05-08 R/connect.R
Add extended examples for metadata functions.

@@ -108,9 +108,23 @@ cites_shipments <- function() {
 #' @aliases metadata cites_metadata
 #' @examples
 #' if (cites_status()) {
+#'   library(dplyr)
+#'
+#'   # Look at the field definitions for cites_shipments()
 #'   cites_metadata()
-#'   cites_codes()
-#'   cites_parties()
+
+#'   # View the codes used for purpose of shipment
+#'   cites_codes() %>%
+#'    filter(field == "Purpose")
+#'
+#'   # See the most recent countries to join CITES
+#'   cites_parties() %>%
+#'     arrange(desc(date)) %>%
+#'     head(10)
+#'
+#'   # See countries or locations with non-standaard or outdated ISO codes
+#'   cites_parties() %>%
+#'     filter(former_code | non_ISO_code)
 #'
 #'   # For remote connections to these tables,
 #'   # access the database directly:

### 2019-05-08 inst/response_to_reviewers.md
More elaborate examples and help text.

@@
 # Response to reviewers
+
+Thanks to both reviewers for their comments and issuees
+
+## Changes implemented
 
@@
   should trigger disk cleanup and avoid doubling database size
+- We have modified out linter tests to avoid the false positives shown in
+  https://github.com/ecohealthalliance/citesdb/issues/2
+- We have renamed `cites_status()` to `cites_check_status()` to be less generic
+
+## Changes not implemented / justifications
 
@@
   for interacting with the RStudio connection pane, which can not be
-  straightforwardly testeed. Other code coverage is 
+  straightforwardly tested. Other code coverage is 
   [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)

### 2019-05-08 R/connect.R

@@ -31,12 +31,14 @@ cites_check_status <- function() {
 #' @export
 #'
 #' @examples
-#' \donttest{
-#' \dontrun{
-#' library(DBI)
-#' dbListTables(cites_db())
-#' parties <- dbReadTable(cites_db(), "cites_parties")
-#' }
+#' if (cites_status()) {
+#'   library(DBI)
+#'   dbListTables(cites_db())
+#'   parties <- dbReadTable(cites_db(), "cites_parties")
+#'   dbGetQuery(
+#'    cites_db(),
+#'    'SELECT "Taxon", "Importer" FROM cites_shipments WHERE "Year" = 1976 LIMIT 100;'
+#'    )
 #' }
 cites_db <- function(dbdir = cites_path()) {
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
@@ -72,13 +74,33 @@ cites_db <- function(dbdir = cites_path()) {
 
 #' CITES shipment data
 #'
-#' Returns a remote table with all CITES shipment data. Requires the dplyr and dbplyr packages.
+#' Returns a remote database table with all CITES shipment data.  This is the
+#' bulk of the data in the package and constitutes > 2 million records.  Loading
+#' the whole table into R via the [dplyr::colleect()] command will use over
+#' 3 GB of RAM, so you may want to pre-process data in the database, as in
+#' the examples below.
+#'
 #' @return A dplyr remote tibble ([dplyr::tbl()])
 #' @export
 #'
 #' @examples
 #' if (cites_status()) {
-#'   cites_shipments()
+#'   library(dplyr)
+#'
+#'   # Examine number of records per year.
+#'   cites_shipments() %>%
+#'     group_by(Year) %>%
+#'     summarize(n_records = n()) %>%
+#'     arrange(desc(Year)) %>%
+#'     collect()
+#'
+#'   # See what Pangolin shipments went to what countries in 1990
+#'    cites_shipments() %>%
+#'      filter(Order == "Pholidota", Year == 1990) %>%
+#'      count(Year, Importer, Term) %>% collect() %>%
+#'      left_join(select(cites_parties(), country, code),
+#'                by = c("Importer"="code"))
+#'
 #' }
 #' @importFrom dplyr tbl
 cites_shipments <- function() {

### 2019-05-08 R/download.R
Fix linting error

@@ -157,9 +157,10 @@ load_citesdb_metadata <- function() {
   tblnames <- tools::file_path_sans_ext(basename(tsvs))
   for (i in seq_along(tsvs)) {
     suppressMessages(dbWriteTable(cites_db(), tblnames[i],
-                                  read.table(tsvs[i],
-                                             stringsAsFactors = FALSE, sep = "\t",
-                                             header = TRUE, quote = "\""
+                                  read.table(
+                                    tsvs[i],
+                                    stringsAsFactors = FALSE, sep = "\t",
+                                    header = TRUE, quote = "\""
                                   ),
                                   overwrite = TRUE
     ))

### 2019-05-08 inst/response_to_reviewers.md
Update wordlist and response to reviewers

@@
 
-Thanks to both reviewers for their comments and issuees
+Thanks to both reviewers for their comments and issues
 
@@
 - We have renamed `cites_status()` to `cites_check_status()` to be less generic
+- We have made more elaborate help and examples for the `cites_db()`, 
+  `cites_shipments()`, and metadata functions to illustrate their use and
+  distinguish between **dplyr** and **DBI**-based workflows.
 
@@
   and all major funcionality is tested.
+
+We have not yet been able to recreate the error in
+https://github.com/ecohealthalliance/citesdb/issues/4, but have made some
+changes that may resolve it.  Can reviewers test this and report back in that
+issue?
+
+A note - we have heard from the maintainers of MonetDBLite that it will not be
+returning to CRAN, but they are working on a successor embeedddeed database package that
+will replace it and go to CRAN later this year. So for now we will host this
+package on GitHub and replace the database back-end and send to CRAN when
+the successor is ready.

### 2019-05-08 R/connect.R
Fix typo

@@ -76,7 +76,7 @@ cites_db <- function(dbdir = cites_path()) {
 #'
 #' Returns a remote database table with all CITES shipment data.  This is the
 #' bulk of the data in the package and constitutes > 2 million records.  Loading
-#' the whole table into R via the [dplyr::colleect()] command will use over
+#' the whole table into R via the [dplyr::collect()] command will use over
 #' 3 GB of RAM, so you may want to pre-process data in the database, as in
 #' the examples below.
 #'

### 2019-05-08 inst/response_to_reviewers.md
Tweak response to reviewers

@@
 A note - we have heard from the maintainers of MonetDBLite that it will not be
-returning to CRAN, but they are working on a successor embeedddeed database package that
-will replace it and go to CRAN later this year. So for now we will host this
-package on GitHub and replace the database back-end and send to CRAN when
-the successor is ready.
+returning to CRAN (it's current iteration fails on R-devel), but they are
+working on a successor embeedddeed database package that will replace it and go
+to CRAN later this year. So for now we will host this package on GitHub and
+replace the database back-end and send to CRAN when the successor is ready.

### 2019-05-08 inst/response_to_reviewers.md
Tweak response to reviewers (2)

@@
   for interacting with the RStudio connection pane, which can not be
-  straightforwardly tested. Other code coverage is 
+  tested other than in an interactive session. Other code coverage is 
   [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)

### 2019-05-08 inst/response_to_reviewers.md
Fix typo

@@
 returning to CRAN (it's current iteration fails on R-devel), but they are
-working on a successor embeedddeed database package that will replace it and go
+working on a successor embedded database package that will replace it and go
 to CRAN later this year. So for now we will host this package on GitHub and

### 2019-05-08 inst/response_to_reviewers.md
More text on testing

@@
   tested other than in an interactive session. Other code coverage is 
-  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R)
-  and all major funcionality is tested.
+  [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R).
+  We believe that all important functionality is tested, including important
+  edge-cases and conditions not reflected in the coverage statistic, such as
+  error handling in multiple sessions and changing up upstream data sources.
 

### 2019-05-08 inst/response_to_reviewers.md
Various edits to review fixes

@@
 
-Thanks to both reviewers for their comments and issues
+Thanks to both reviewers for their comments and issues that have 
+helped us to make a more user-friendly package.
 
@@
 
-- We have changed the installation method to use `remotes` (which has lower
-  dependencies than `devtools`).
-- We have added disk location to `cites_status`
+- We have changed the installation method in the `README` to use **remotes**, 
+  which has fewer dependencies than **devtools**
+
+- We have added information to the `cites_status` table indicating the location 
+  of the database files on disk
+
 - We now clear out and disconnect the database tables after updating, which
   should trigger disk cleanup and avoid doubling database size
-- We have modified out linter tests to avoid the false positives shown in
+
+- We have modified our linter tests to avoid the false positives shown in
   https://github.com/ecohealthalliance/citesdb/issues/2
-- We have renamed `cites_status()` to `cites_check_status()` to be less generic
+
+- We have renamed the internal function `check_status()` to `cites_check_status()`
+  to be less generic
+
 - We have made more elaborate help and examples for the `cites_db()`, 
@@
 
-- We have opted not to use **glue** but stick with `paste` in the interest of
-  limiting dependencies.  This is a trade-off but we believe a minor one.
+- We have opted not to use the **glue** package and instead stick with the base
+  R `paste()` functions in the interest of limiting dependencies. We believe
+  this is a minor trade-off.
+
 - The low test coverage shown in https://github.com/ecohealthalliance/citesdb/issues/4
-  is due to tests skippeed on CRAN.  Setting the environment variable to 
+  is due to tests skipped on CRAN. Setting the environment variable to 
   `NOT_CRAN=true` shows that our test coverage is 65%. This is lower than
-  typical, but as we note above, this is largely due to the verbose code
-  for interacting with the RStudio connection pane, which can not be
-  tested other than in an interactive session. Other code coverage is 
+  typical, but as we note in the issue above, this is largely due to the 
+  verbose code for interacting with the RStudio connection pane, which cannot be
+  tested except in an interactive session. Other code coverage is 
   [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R).
@@
 We have not yet been able to recreate the error in
-https://github.com/ecohealthalliance/citesdb/issues/4, but have made some
+https://github.com/ecohealthalliance/citesdb/issues/1, but have made some
 changes that may resolve it. Can reviewers test this and report back in that
@@
 
-A note - we have heard from the maintainers of MonetDBLite that it will not be
-returning to CRAN (it's current iteration fails on R-devel), but they are
+A final note: we have heard from the maintainers of **MonetDBLite** that it will 
+not be returning to CRAN (its current iteration fails on R-devel), but they are
 working on a successor embedded database package that will replace it and go
-to CRAN later this year. So for now we will host this package on GitHub and
-replace the database back-end and send to CRAN when the successor is ready.
+to CRAN later this year. So, for now, we will host this package on GitHub and
+replace the database back-end and send to CRAN when the successor package is ready.

### 2019-05-08 README.Rmd
Various edits to review fixes

@@
 
-When you first load the package you will see a message like this:
+When you first load the package, you will see a message like this:
 
@@
 
-Once you fetch the data you can connect to the database with the `cites_db()` command.
-You can use the `cites_shipments()` command to load a remote `tibble` that is backed by the database but not loaded into R.
-You can use this to analyze CITES data without ever loading it into memory, then gather your results with the `dplyr` function `collect()`.
+Once you fetch the data, you can connect to the database with the `cites_db()` command.
+The `cites_shipments()` command loads a remote `tibble` that is backed by the database but is not loaded into R.
+You can use this command to analyze CITES data without ever loading it into memory, gathering your results with the `dplyr` function `collect()`.
 For example:
@@
 
-The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making such analyses quite snappy even on such large data using normal desktops and laptops.
+The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making analyses on such large data quite snappy using normal desktops and laptops.
 Here's the timing of the above query, which processes over 20 million records:
@@
 
-If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from original source at <https://trade.cites.org/>.
+If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from the original source at <https://trade.cites.org/>.
 
@@
 citesdb::cites_disconnect()
-MonetDBLite::monetdblite_shutdown()
 ```

### 2019-05-08 R/connect.R

@@ -33,8 +33,11 @@ cites_check_status <- function() {
 #' @examples
 #' if (cites_status()) {
 #'   library(DBI)
+#'
 #'   dbListTables(cites_db())
+#'
 #'   parties <- dbReadTable(cites_db(), "cites_parties")
+#'
 #'   dbGetQuery(
 #'    cites_db(),
 #'    'SELECT "Taxon", "Importer" FROM cites_shipments WHERE "Year" = 1976 LIMIT 100;'
@@ -75,31 +78,32 @@ cites_db <- function(dbdir = cites_path()) {
 #' CITES shipment data
 #'
 #' Returns a remote database table with all CITES shipment data.  This is the
-#' bulk of the data in the package and constitutes > 2 million records.  Loading
+#' bulk of the data in the package and constitutes > 20 million records.  Loading
 #' the whole table into R via the [dplyr::collect()] command will use over
 #' 3 GB of RAM, so you may want to pre-process data in the database, as in
 #' the examples below.
 #'
-#' @return A dplyr remote tibble ([dplyr::tbl()])
+#' @return A **dplyr** remote tibble ([dplyr::tbl()])
 #' @export
 #'
 #' @examples
 #' if (cites_status()) {
 #'   library(dplyr)
 #'
-#'   # Examine number of records per year.
+#'   # See the number of CITES shipment records per year
 #'   cites_shipments() %>%
 #'     group_by(Year) %>%
 #'     summarize(n_records = n()) %>%
 #'     arrange(desc(Year)) %>%
 #'     collect()
 #'
-#'   # See what Pangolin shipments went to what countries in 1990
+#'   # See what pangolin shipments went to which countries in 1990
 #'    cites_shipments() %>%
 #'      filter(Order == "Pholidota", Year == 1990) %>%
-#'      count(Year, Importer, Term) %>% collect() %>%
+#'      count(Year, Importer, Term) %>%
+#'      collect() %>%
 #'      left_join(select(cites_parties(), country, code),
-#'                by = c("Importer"="code"))
+#'                by = c("Importer" = "code"))
 #'
 #' }
 #' @importFrom dplyr tbl
@@ -132,10 +136,10 @@ cites_shipments <- function() {
 #' if (cites_status()) {
 #'   library(dplyr)
 #'
-#'   # Look at the field definitions for cites_shipments()
+#'   # See the field definitions for cites_shipments()
 #'   cites_metadata()
-
[…]

### 2019-05-21 inst/response_to_reviewers.md
Update response to reviewers

@@
 - We have changed the installation method in the `README` to use **remotes**, 
-  which has fewer dependencies than **devtools**
+  which has fewer dependencies than **devtools**.
 
 - We have added information to the `cites_status` table indicating the location 
-  of the database files on disk
+  of the database files on disk.
 
 - We now clear out and disconnect the database tables after updating, which
-  should trigger disk cleanup and avoid doubling database size
+  should trigger disk cleanup and avoid doubling database size.
+
+- We have solved the error that was preventing package-building in 
+  https://github.com/ecohealthalliance/citesdb/issues/1, which had its origins
+  in missing a token for using the **rcites** package.  We now cache this
+  information to avoid having to make remote calls in vignette-buildng.
 
 - We have modified our linter tests to avoid the false positives shown in
-  https://github.com/ecohealthalliance/citesdb/issues/2
+  https://github.com/ecohealthalliance/citesdb/issues/2 .
 
 - We have renamed the internal function `check_status()` to `cites_check_status()`
-  to be less generic
+  to be less generic.
 
@@
   `cites_shipments()`, and metadata functions to illustrate their use and
-  distinguish between **dplyr** and **DBI**-based workflows.
+  distinguish between **dplyr**- and **DBI**-based workflows.
 
@@
 
-We have not yet been able to recreate the error in
-https://github.com/ecohealthalliance/citesdb/issues/1, but have made some
-changes that may resolve it. Can reviewers test this and report back in that
-issue?
-
-A final note: we have heard from the maintainers of **MonetDBLite** that it will 
+A final note: we have leaened from the maintainers of **MonetDBLite** that it will 
 not be returning to CRAN (its current iteration fails on R-devel), but they are

### 2019-05-21 inst/response_to_reviewers.md
Minor tweaks to response to reviewers

@@
   https://github.com/ecohealthalliance/citesdb/issues/1, which had its origins
-  in missing a token for using the **rcites** package.  We now cache this
-  information to avoid having to make remote calls in vignette-buildng.
+  in a missing token for use with the **rcites** package. We now cache this
+  information to avoid having to make remote calls in vignette-building.
 
 - We have modified our linter tests to avoid the false positives shown in
-  https://github.com/ecohealthalliance/citesdb/issues/2 .
+  https://github.com/ecohealthalliance/citesdb/issues/2.
 
@@
   [greater than 90%](https://codecov.io/gh/ecohealthalliance/citesdb/tree/master/R).
-  We believe that all important functionality is tested, including important
+  We believe that all core functionality is tested, including important
   edge-cases and conditions not reflected in the coverage statistic, such as
@@
 
-A final note: we have leaened from the maintainers of **MonetDBLite** that it will 
+A final note: we have learned from the maintainers of **MonetDBLite** that it will 
 not be returning to CRAN (its current iteration fails on R-devel), but they are

### 2019-05-22 R/connect.R
Fix cites_disconect typos

@@ -62,7 +62,7 @@ cites_db <- function(dbdir = cites_path()) {
       if (grepl("(Database lock|bad rolemask)", e)) {
         stop(paste(
           "Local citesdb database is locked by another R session.\n",
-          "Try closing or running cites_disconect() in that session."
+          "Try closing or running cites_disconnect() in that session."
         ),
         call. = FALSE
         )

### 2019-05-22 R/connect.R
More aggressive conneection management
-  Override MonetDB lock when connecting
-  Do not connect to pane inside a devtools::check()
   command.

@@ -55,6 +55,7 @@ cites_db <- function(dbdir = cites_path()) {
 
   tryCatch(
     {
+      unlink(file.path(dbdir, ".gdk_lock"))
       db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir)
     },
     error = function(e) {
@@ -193,6 +194,7 @@ cites_disconnect_ <- function(environment = cites_cache) { # nolint
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
     DBI::dbDisconnect(db, shutdown = TRUE)
+    MonetDBLite::monetdblite_shutdown()
   }
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {
@@ -202,3 +204,4 @@ cites_disconnect_ <- function(environment = cites_cache) { # nolint
 
 cites_cache <- new.env()
 reg.finalizer(cites_cache, cites_disconnect_, onexit = TRUE)
+

### 2019-05-22 R/connection-pane.R

@@ -29,7 +29,7 @@ sql_action <- function() {
 #' if (!is.null(getOption("connectionObserver"))) cites_pane()
 cites_pane <- function() {
   observer <- getOption("connectionObserver")
-  if (!is.null(observer)) {
+  if (!is.null(observer) && interactive()) {
     observer$connectionOpened(
       type = "CITESDB",
       host = "citesdb",

### 2019-05-22 R/onattach.R

@@ -1,6 +1,14 @@
+in_chk <- function() {
+  any(
+    grepl("check",
+          sapply(sys.calls(), function(a) paste(deparse(a), collapse = "\n"))
+    )
+  )
+}
+
 .onAttach <- function(libname, pkgname) {
   MonetDBLite::monetdblite_shutdown()
-  if (interactive() && Sys.getenv("RSTUDIO") == "1") {
+  if (interactive() && Sys.getenv("RSTUDIO") == "1"  && !in_chk()) {
     cites_pane()
   }
   if (interactive()) cites_status()
@@ -40,14 +48,14 @@ cites_db_delete <- function() {
 #' cites_status()
 cites_status <- function(verbose = TRUE) {
   if (dbExistsTable(cites_db(), "cites_shipments") &&
-    dbExistsTable(cites_db(), "cites_status")) {
+      dbExistsTable(cites_db(), "cites_status")) {
     status <- DBI::dbReadTable(cites_db(), "cites_status")
     status_msg <-
       paste0(
         "CITES database status:\n",
         paste0(toTitleCase(gsub("_", " ", names(status))),
-          ": ", as.matrix(status),
-          collapse = "\n"
+               ": ", as.matrix(status),
+               collapse = "\n"
         )
       )
     out <- TRUE

### 2019-05-22 R/connect.R
Remove gc() cleanup, previously used for debug

@@ -55,7 +55,6 @@ cites_db <- function(dbdir = cites_path()) {
 
   tryCatch(
     {
-      gc(verbose = FALSE)
       db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir)
     },
     error = function(e) {
@@ -193,7 +192,6 @@ cites_disconnect <- function() {
 cites_disconnect_ <- function(environment = cites_cache) { # nolint
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
-    gc(verbose = FALSE)
     DBI::dbDisconnect(db, shutdown = TRUE)
   }
   observer <- getOption("connectionObserver")

### 2019-05-22 README.Rmd
Add more awknowledgements of WCMC
Note need for package build tools.

@@
 This data consists of over 40 years and 20 million records of reported shipments of wildlife and wildlife products subject to oversight under the [Convention on International Trade in Endangered Species of Wild Fauna and Flora](https://www.cites.org).
+The source data are maintained by the [UN Environment World Conservation Monitoring Centre](https://www.unep-wcmc.org/).
 
@@
 ```{r install_me, eval = FALSE}
-remotes::install_github("ecohealthalliance/citesdb")
+devtools::install_github("ecohealthalliance/citesdb")
 ```
+
+Note that since **citesdb** installs a source dependency from GitHub, you
+will need [package build tools](http://stat545.com/packages01_system-prep.html).
 

### 2019-05-22 inst/response_to_reviewers.md
Update response to reviewers

@@
 
-- We have changed the installation method in the `README` to use **remotes**, 
-  which has fewer dependencies than **devtools**.
+- We have changed the installation method in the `README` to use **devtools**.
 
@@
 
-- We have solved the error that was preventing package-building in 
+- Thanks for the reviewers for sticking with us through a long and merry chase
+  of a bug that was preventing package-building in 
   https://github.com/ecohealthalliance/citesdb/issues/1, which had its origins
-  in a missing token for use with the **rcites** package. We now cache this
-  information to avoid having to make remote calls in vignette-building.
+  in a missing token for use with the **rcites** package as well as a low-level
+  database lock issue. We now cache this the **rcites** information to avoid
+  having to make remote calls in vignette-building, and also have resolved the
+  DB locking conflict.
 
@@
 replace the database back-end and send to CRAN when the successor package is ready.
+We've added a note to the README showing that users need package build tools for
+the current version.

### 2019-05-29 inst/pangolin-anim/app-data-process.R
Add pangolin app in `inst/`

@@ -1,0 +1,51 @@
+library(citesdb)
+library(tidyverse)
+library(CoordinateCleaner)
+library(countrycode)
+library(sf)
+
+pangolin_shipments <- cites_shipments() %>%
+  filter(Order == "Pholidota", Purpose %in% c(NA_character_, "T")) %>%
+  select(-Purpose, -Import.permit.RandomID, -Export.permit.RandomID, -Origin.permit.RandomID, -Class, -Order, -Family, -Genus, -Reporter.type ) %>%
+  collect()
+cites_disconnect()
+
+cref <- CoordinateCleaner::countryref %>%
+  filter(type == "country") %>%
+  distinct(iso2, .keep_all = TRUE) %>%
+  select(iso2, centroid.lon, centroid.lat)
+
+pangolin_shipments <- pangolin_shipments %>%
+  mutate(start = coalesce(Origin, Exporter)) %>%
+  left_join(cref, by = c("start" = "iso2")) %>%
+  rename(start.lon = centroid.lon, start.lat = centroid.lat) %>%
+  left_join(cref, by = c("Importer" = "iso2")) %>%
+  rename(end.lon = centroid.lon, end.lat = centroid.lat) %>%
+  filter(Importer != start) %>%
+  filter(!is.na(start.lon), !is.na(end.lon), !is.na(start.lat), !is.na(end.lat)) %>%
+  mutate(id = seq_len(n()))
+
+
+dat <- pangolin_shipments %>%
+  janitor::clean_names() %>%
+  mutate(
+    # reporting_country = ifelse(reporter_type == "I", importer, exporter),
+    # reporting_country  = countrycode(sourcevar = reporting_country,
+    #                                  origin = "iso2c",
+    #                                  destination = "country.name"),
+    start2 = countrycode(sourcevar = start,
+                         origin = "iso2c",
+                         destination = "country.name"),
+    end = countrycode(sourcevar = importer,
+                      origin = "iso2c",
+                      destination = "country.name")) %>%
+  select(year, start2, end, term, start_lon, start_lat, end_lon, end_lat) %>%
+  rename(start = start2)
+
+
+dat2 <- dat %>%
+  group_by_all() %>%
+  mutate(n = n()) %>%
+  ungroup()
+
+write_rds(dat2, here::here("inst", "pangolin-anim", "pangolin_dat.rds"))

### 2019-05-29 inst/pangolin-anim/app.R

@@ -1,0 +1,80 @@
+library(shiny)
+library(tidyverse)
+library(echarts4r)
+library(RColorBrewer)
+
+dat <- read_rds("pangolin_dat.rds") %>%
+    arrange(year, n)
+
+ui <- fluidPage(
+
+    sidebarLayout(
+        sidebarPanel(width = 3,
+                     sliderInput("year",
+                                 "Year",
+                                 min = min(dat$year, na.rm = TRUE),
+                                 max = max(dat$year, na.rm = TRUE),
+                                 value = 2017,
+                                 sep = "",
+                                 step = 1)
+
+        ),
+        mainPanel(
+            tabsetPanel(
+                tabPanel("Connectivity Map - echarts4r",
+                         fluidRow(
+                             column(12, echarts4rOutput("map"))
+                         )
+                )
+            )
+        )
+    )
+)
+
+server <- function(input, output) {
+
+    #https://ecomfe.github.io/echarts-examples/public/editor.html?c=geo-lines
+    # echarts4r map
+    output$map <- renderEcharts4r({
+
+        dat2 = dat %>%
+            filter(year == input$year
+            ) %>%
+            group_by(term)
+
+        pal <-  scales::hue_pal()(n_distinct(dat2$term))
+
+        my_scale <- function(x) scales::rescale(x, to = c(3, 10))
+
+        # note cannot add bind to e_lines...can fork repo and try?
+        dat2 %>%
+            e_charts(start_lon) %>%
+            e_geo(roam = TRUE, label = list(emphasis = list(show = FALSE)),
+                  itemStyle = list(normal = list(areaColor = '#323c48', borderColor = '#404a59'),
+                                   emphasis = list(areaColor = '#2a333d', borderColor = '#2a333d')
+                                   )
+            ) %>%
+            e_color(background = "black") %>%
+            e_theme("chalk") %>%
+            e_lines(start_lon, start_lat, end_lon, end_lat, silent = TRUE,
+                    effect = list(show=T, period=3, trailLength=0, symbolSize=3),
+                    lineStyle = list(normal = list(curveness = -.20, width = 1))) %>%
+            e_effect_scatter(start_lat, coord_system = "geo",
+                             size = n,
+                             scale = my_scale,
+                             #symbol_size = 1,
+                             bind = start,
+                             silent = FALSE) %>%
+            #e_tooltip(trigger = "item") %>%
+            e_tooltip(trigger = "item",
[…]

### 2019-05-29 README.Rmd
Change repo linkss to ropensci org

@@
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2630836.svg){data-external="1"}](https://doi.org/10.5281/zenodo.2630836)
-[![CircleCI](https://circleci.com/gh/ecohealthalliance/citesdb/tree/master.svg?style=shield){data-external="1"}](https://circleci.com/gh/ecohealthalliance/citesdb)
+[![CircleCI](https://circleci.com/gh/ropensci/citesdb/tree/master.svg?style=shield){data-external="1"}](https://circleci.com/gh/ropensci/citesdb)
 [![](https://badges.ropensci.org/292_status.svg)](https://github.com/ropensci/software-review/issues/292)
-[![codecov](https://codecov.io/gh/ecohealthalliance/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ecohealthalliance/citesdb)
+[![codecov](https://codecov.io/gh/ropensci/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ropensci/citesdb)
 [![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)
@@
 ```{r install_me, eval = FALSE}
-devtools::install_github("ecohealthalliance/citesdb")
+devtools::install_github("ropensci/citesdb")
 ```
@@
 
-If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ecohealthalliance/citesdb/releases), or as a `.zip` file of many CSV files from the original source at <https://trade.cites.org/>.
+If you don't need any of the bells and whistles of this package, you can download the raw data as a single compressed TSV file from the [releases page](https://github.com/ropensci/citesdb/releases), or as a `.zip` file of many CSV files from the original source at <https://trade.cites.org/>.
 
@@
 Great!
-Please take a look at the [contributing guidelines](https://github.com/ecohealthalliance/citesdb/blob/master/.github/CONTRIBUTING.md) before filing an issue or pull request.
+Please take a look at the [contributing guidelines](https://github.com/ropensci/citesdb/blob/master/.github/CONTRIBUTING.md) before filing an issue or pull request.
 
-Please note that this project is released with a [Contributor Code of Conduct](https://github.com/ecohealthalliance/citesdb/blob/master/.github/CODE_OF_CONDUCT.md).
+Please note that this project is released with a [Contributor Code of Conduct](https://github.com/ropensci/citesdb/blob/master/.github/CODE_OF_CONDUCT.md).
 By participating in this project you agree to abide by its terms.
 
-[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ecohealthalliance/citesdb/master/vignettes/figures/eha-footer.png){data-external="1"}](https://www.ecohealthalliance.org/)
+[![Created by EcoHealth Alliance](https://raw.githubusercontent.com/ropensci/citesdb/master/vignettes/figures/eha-footer.png){data-external="1"}](https://www.ecohealthalliance.org/)

### 2019-05-29 R/download.R

@@ -9,7 +9,7 @@
 #'
 #' @param tag What release tag of data to download. Defaults to the most recent.
 #' Releases are expected to come twice per year. See all releases at
-#' <https://github.com/ecohealthalliance/citesdb/releases>.
+#' <https://github.com/ropensci/citesdb/releases>.
 #' @param destdir Where to download the compressed file.
 #' @param cleanup Whether to delete the compressed file after loading into the database.
 #' @param verbose Whether to display messages and download progress
@@ -29,7 +29,7 @@
 cites_db_download <- function(tag = NULL, destdir = tempdir(),
                               cleanup = TRUE, verbose = interactive()) {
   if (verbose) message("Downloading data...\n")
-  zfile <- get_gh_release_file("ecohealthalliance/citesdb",
+  zfile <- get_gh_release_file("ropensci/citesdb",
     tag_name = tag,
     destdir = destdir, verbose = verbose
   )

### 2019-05-29 data-raw/process_trade_db.R

@@ -65,7 +65,7 @@ file_delete(
 
 # Release the compressed data
 datastorr::github_release_info(
-  "ecohealthalliance/citesdb",
+  "ropensci/citesdb",
   read = read_tsv,
   filename = "cites_trade_db.tsv.bz2"
 ) %>%

### 2019-05-29 README.Rmd
Change status from WIP to Active

@@
 [![codecov](https://codecov.io/gh/ropensci/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ropensci/citesdb)
-[![Project Status: WIP - Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](http://www.repostatus.org/badges/latest/wip.svg){data-external="1"}](http://www.repostatus.org/#wip)
+[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
 

### 2019-05-31 README.Rmd
Minor documentation fixes, reknit README
Update bibliographical info to JOSS
- Addd badge and rebuild README
- Bump cache to build site with updated pkgdown

@@
 
+<!-- badges: start -->
+
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
+[![Published in the Journal of Open Source Software](http://joss.theoj.org/papers/10.21105/joss.01483/status.svg){data-external="1"}](https://doi.org/10.21105/joss.01483)
 [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2630836.svg){data-external="1"}](https://doi.org/10.5281/zenodo.2630836)
@@
 [![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
+
+<!-- badges: end -->
 
@@
 It will expand to over 1 GB in the local database.
-During the download and database building up to 3.5 GB of disk space may be used temporarily.
+During the download and database building, up to 3.5 GB of disk space may be used temporarily.
 

### 2019-06-28 README.Rmd
Add rOpenSci Badge

@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
+[![rOpensci\_Badge](https://badges.ropensci.org/292_status.svg)](https://github.com/ropensci/onboarding/issues/292)
 [![Published in the Journal of Open Source Software](http://joss.theoj.org/papers/10.21105/joss.01483/status.svg){data-external="1"}](https://doi.org/10.21105/joss.01483)

### 2019-07-31 README.Rmd
Removing duplicate badge

--- 2019-05-31 README.Rmd
@@
 [![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg){data-external="1"}](https://opensource.org/licenses/MIT)
+[![rOpensci\_Badge](https://badges.ropensci.org/292_status.svg)](https://github.com/ropensci/software-review/issues/292)
 [![Published in the Journal of Open Source Software](http://joss.theoj.org/papers/10.21105/joss.01483/status.svg){data-external="1"}](https://doi.org/10.21105/joss.01483)
@@
 [![CircleCI](https://circleci.com/gh/ropensci/citesdb/tree/master.svg?style=shield){data-external="1"}](https://circleci.com/gh/ropensci/citesdb)
-[![](https://badges.ropensci.org/292_status.svg)](https://github.com/ropensci/software-review/issues/292)
 [![codecov](https://codecov.io/gh/ropensci/citesdb/branch/master/graph/badge.svg){data-external="1"}](https://codecov.io/gh/ropensci/citesdb)

### 2020-09-28 R/citesdb-package.R
monetdblite -> duckdb

@@ -8,7 +8,7 @@
 #' This package allows you to download the full database of historic transactions
 #' (commonly called the [CITES Trade Database](https://trade.cites.org/))
 #' to analyze locally on your computer. Because of the large size of this data,
-#' the **citesdb** package stores it as an on-disk [MonetDB](https://www.monetdb.org/)
+#' the **citesdb** package stores it as an on-disk [DuckDB](https://duckdb.org/)
 #' that can be queried without loading into RAM.
 #'
 #' @author Noam Ross and Evan A. Eskew, [EcoHealth Alliance](https://www.ecohealthalliance.org/)

### 2020-09-28 R/download.R

@@ -2,7 +2,7 @@
 #'
 #' This command downloads the CITES shipments database and populates a local
 #' database. The download is large (~158 MB), and the database will be over
-#' 1 GB on disk. During import over 3.5 GB of disk space may be used temporarily.
+#' 4.7 GB on disk. During import over 5 GB of disk space may be used temporarily.
 #'
 #' The database is stored by default under [rappdirs::user_data_dir()], or its
 #' location can be set with the environment variable `CITES_DB_DIR`.
@@ -19,6 +19,7 @@
 #' @importFrom DBI dbRemoveTable dbExistsTable dbCreateTable dbExecute
 #'   dbWriteTable dbListTables
 #' @importFrom R.utils gunzip
+#' @importFrom dplyr %>%
 #'
 #' @examples
 #' \donttest{
@@ -44,15 +45,16 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
 
   cites_disconnect()
   tblname <- "cites_shipments"
+
   dbCreateTable(cites_db(), tblname, fields = cites_field_types)
 
   suppressMessages(
     dbExecute(
       cites_db(),
       paste0(
-        "COPY OFFSET 2 INTO ", tblname, " FROM '",
+        "COPY ", tblname, " FROM '",
         temp_tsv,
-        "' USING DELIMITERS '\t','\n','\"' NULL as 'NA'"
+        "' ( DELIMITER '\t', HEADER )"
       )
     )
   )

### 2020-09-28 R/onattach.R

@@ -7,7 +7,7 @@ in_chk <- function() {
 }
 
 .onAttach <- function(libname, pkgname) {  #nolint
-  MonetDBLite::monetdblite_shutdown()
+  duckdb::duckdb_shutdown(duckdb::duckdb())
   if (interactive() && Sys.getenv("RSTUDIO") == "1"  && !in_chk()) {
     cites_pane()
   }

### 2020-09-28 tests/testthat/test-03-connection.R

@@ -14,7 +14,7 @@ test_that("Disconnetion works", {
     success <- callr::r(function() {
       options(CITES_DB_DIR = "localdb")
       con <- DBI::dbConnect(
-        MonetDBLite::MonetDBLite(),
+        duckdb::duckdb(),
         getOption("CITES_DB_DIR")
       )
       out <- inherits(con, "MonetDBEmbeddedConnection")

### 2020-09-28 R/download.R
Code NAs correctly

@@ -54,7 +54,7 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
       paste0(
         "COPY ", tblname, " FROM '",
         temp_tsv,
-        "' ( DELIMITER '\t', HEADER )"
+        "' ( DELIMITER '\t', HEADER 1, NULL 'NA' )"
       )
     )
   )

### 2020-09-28 README.Rmd
monetdblite -> duckdb (#10)
* monetdblite -> duckdb
* Fix connection pane to work with duckdb
* Code NAs correctly

Co-authored-by: Noam Ross <ross@ecohealthalliance.org>

@@
 citesdb::cites_disconnect()
-MonetDBLite::monetdblite_shutdown()
+duckdb::duckdb_shutdown()
 ```

### 2020-09-28 R/connect.R

@@ -17,7 +17,7 @@ cites_check_status <- function() {
 #' The local CITES database
 #'
 #' Returns a connection to the local CITES database. This is a DBI-compliant
-#' [MonetDBLite::MonetDBLite()] database connection. When using **dplyr**-based
+#' [duckdb::duckdb()] database connection. When using **dplyr**-based
 #' workflows, one typically accesses tables with functions such as
 #' [cites_shipments()], but this function lets one interact with the database
 #' directly via SQL.
@@ -25,9 +25,9 @@ cites_check_status <- function() {
 #' @param dbdir The location of the database on disk. Defaults to
 #' `citesdb` under [rappdirs::user_data_dir()], or the environment variable `CITES_DB_DIR`.
 #'
-#' @return A MonetDBLite DBI connection
+#' @return A DuckDB DBI connection
 #' @importFrom DBI dbIsValid dbConnect
-#' @importFrom MonetDBLite MonetDBLite
+#' @importFrom duckdb duckdb
 #' @export
 #'
 #' @examples
@@ -55,7 +55,7 @@ cites_db <- function(dbdir = cites_path()) {
 
   tryCatch({
       unlink(file.path(dbdir, ".gdk_lock"))
-      db <- DBI::dbConnect(MonetDBLite::MonetDBLite(), dbname = dbdir)
+      db <- DBI::dbConnect(duckdb::duckdb(), paste0(dbdir, "/citesdb.duckdb"))
     },
     error = function(e) {
       if (grepl("(Database lock|bad rolemask)", e)) {
@@ -193,7 +193,7 @@ cites_disconnect_ <- function(environment = cites_cache) { # nolint
   db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
     DBI::dbDisconnect(db, shutdown = TRUE)
-    MonetDBLite::monetdblite_shutdown()
+    duckdb::duckdb_shutdown(duckdb::duckdb())
   }
   observer <- getOption("connectionObserver")
   if (!is.null(observer)) {

### 2020-09-28 R/connection-pane.R

@@ -51,11 +51,10 @@ cites_pane <- function() {
         )
       },
       listColumns = function(table) {
-        res <- DBI::dbSendQuery(cites_db(),
+        res <- DBI::dbGetQuery(cites_db(),
                                 paste("SELECT * FROM", table, "LIMIT 1"))
-        on.exit(DBI::dbClearResult(res))
         data.frame(
-          name = res@env$info$names, type = res@env$info$types,
+          name = names(res), type = vapply(res, function(x) class(x)[1], character(1)),
           stringsAsFactors = FALSE
         )
       },

### 2020-09-28 R/download.R

@@ -2,7 +2,7 @@
 #'
 #' This command downloads the CITES shipments database and populates a local
 #' database. The download is large (~158 MB), and the database will be over
-#' 1 GB on disk. During import over 3.5 GB of disk space may be used temporarily.
+#' 4.7 GB on disk. During import over 5 GB of disk space may be used temporarily.
 #'
 #' The database is stored by default under [rappdirs::user_data_dir()], or its
 #' location can be set with the environment variable `CITES_DB_DIR`.
@@ -19,6 +19,7 @@
 #' @importFrom DBI dbRemoveTable dbExistsTable dbCreateTable dbExecute
 #'   dbWriteTable dbListTables
 #' @importFrom R.utils gunzip
+#' @importFrom dplyr %>%
 #'
 #' @examples
 #' \donttest{
@@ -44,15 +45,16 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
 
   cites_disconnect()
   tblname <- "cites_shipments"
+
   dbCreateTable(cites_db(), tblname, fields = cites_field_types)
 
   suppressMessages(
     dbExecute(
       cites_db(),
       paste0(
-        "COPY OFFSET 2 INTO ", tblname, " FROM '",
+        "COPY ", tblname, " FROM '",
         temp_tsv,
-        "' USING DELIMITERS '\t','\n','\"' NULL as 'NA'"
+        "' ( DELIMITER '\t', HEADER 1, NULL 'NA' )"
       )
     )
   )

### 2020-09-28 data-raw/process_trade_db.R
Tweaks to update data to 2020.1

@@ -18,9 +18,10 @@ unzip(here("data-raw", "Trade_database.zip"),
 )
 
 tdb <- map_dfr(
-  dir_ls(here("data-raw"), regexp = "^trade_db_\\d+\\.csv$"),
+  list.files(here("data-raw"), pattern = "^trade_db_\\d+\\.csv$", full.names = TRUE),
   ~ read_csv(.,
     col_types = cols(
+      Id = col_character(),
       Year = col_integer(),
       Appendix = col_character(),
       Taxon = col_character(),
@@ -46,7 +47,7 @@ tdb <- map_dfr(
 
 # Removing invalid records to be fixed if possible in the future
 tdb <- tdb %>%
-  filter(!is.na(Year) & Year > 1970 & Year < 2020)
+  filter(!is.na(Year) & Year > 1970 & Year < 2021)
 
 # Arrange values; improves compressibility
 tdb <- tdb %>%
@@ -70,6 +71,6 @@ datastorr::github_release_info(
   filename = "cites_trade_db.tsv.bz2"
 ) %>%
 datastorr::github_release_create(
-  description = "Release of CITES shipment data (v2019.2)",
+  description = "Release of CITES shipment data (v2020.1)",
   target = "master", ignore_dirty = TRUE
 )

### 2020-09-28 data-raw/process_trade_db.R
Update data testing hash

@@ -61,7 +61,7 @@ write_tsv(
 
 # Clean up
 file_delete(
-  dir_ls(here("data-raw"), regexp = "(zip|csv|docx)$")
+  list.files(here("data-raw"), pattern = "(zip|csv|docx)$", full.names = TRUE)
 )
 
 # Release the compressed data

### 2020-10-30 README.Rmd
More tweaks to make duckdb work

@@
 citesdb::cites_disconnect()
-duckdb::duckdb_shutdown()
 ```
@@
 
-The back-end database, [MonetDB](https://monetdb.org), is very fast and powerful, making analyses on such large data quite snappy using normal desktops and laptops.
+The back-end database, [duckdb](https://duckdb.org/), is very fast and powerful, making analyses on such large data quite snappy using normal desktops and laptops.
 Here's the timing of the above query, which processes over 20 million records:

### 2020-10-30 R/connect.R

@@ -43,34 +43,33 @@ cites_check_status <- function() {
 #'    'SELECT "Taxon", "Importer" FROM cites_shipments WHERE "Year" = 1976 LIMIT 100;'
 #'    )
 #' }
-cites_db <- function(dbdir = cites_path()) {
-  db <- mget("cites_db", envir = cites_cache, ifnotfound = NA)[[1]]
+cites_db <- function(dbdir = cites_path(), read_only = TRUE) {
+  db <- mget("cites_db", envir = citesdb:::cites_cache, ifnotfound = NA)[[1]]
   if (inherits(db, "DBIConnection")) {
-    if (DBI::dbIsValid(db)) {
+    if (DBI::dbIsValid(db) && (!read_only && !dbIsReadOnly(db))) {
       return(db)
     }
   }
-  dbname <- dbdir
-  dir.create(dbname, FALSE)
+  dbname <- file.path(dbdir, "citesdb")
+  dir.create(dbdir, FALSE, recursive = TRUE)
 
-  tryCatch({
-      unlink(file.path(dbdir, ".gdk_lock"))
-      db <- DBI::dbConnect(duckdb::duckdb(), paste0(dbdir, "/citesdb.duckdb"))
-    },
-    error = function(e) {
-      if (grepl("(Database lock|bad rolemask)", e)) {
-        stop(paste(
-          "Local citesdb database is locked by another R session.\n",
-          "Try closing or running cites_disconnect() in that session."
-        ),
-        call. = FALSE
-        )
-      } else {
-        stop(e)
-      }
-    },
-    finally = NULL
-  )
+  # tryCatch({
+      db <- DBI::dbConnect(duckdb::duckdb(dbdir=dbname), debug = FALSE, read_only = read_only)
+#
+  #   error = function(e) {
+  #     if (grepl("(Database lock|bad rolemask)", e)) {
+  #       stop(paste(
+  #         "Local citesdb database is locked by another R session.\n",
+  #         "Try closing or running cites_disconnect() in that session."
+  #       ),
+  #       call. = FALSE
+  #       )
+  #     } else {
+  #       stop(e)
+  #     }
+  #   },
+  #   finally = NULL
+  # )
 
   assign("cites_db", db, envir = cites_cache)
   db
@@ -124,7 +123,7 @@ cites_shipments <- function() {
 #' collect the tables into R session memory, rather than returning a remote table.
 #'
 #' This information is drawn from
[…]

### 2020-10-30 R/connection-pane.R

@@ -29,7 +29,7 @@ sql_action <- function() {
 #' if (!is.null(getOption("connectionObserver"))) cites_pane()
 cites_pane <- function() {
   observer <- getOption("connectionObserver")
-  if (!is.null(observer) && interactive()) {
+  if (!is.null(observer) && interactive() && !inherits(try(citesdb::cites_db(),silent = TRUE), "try-error")) {
     observer$connectionOpened(
       type = "CITESDB",
       host = "citesdb",

### 2020-10-30 R/download.R

@@ -1,8 +1,8 @@
 #' Download the CITES database to your local computer
 #'
 #' This command downloads the CITES shipments database and populates a local
-#' database. The download is large (~158 MB), and the database will be over
-#' 4.7 GB on disk. During import over 5 GB of disk space may be used temporarily.
+#' database. The download is large (~158 MB), and the database will be ~5 GB
+#' on disk. During import 10 GB of disk space may be used temporarily.
 #'
 #' The database is stored by default under [rappdirs::user_data_dir()], or its
 #' location can be set with the environment variable `CITES_DB_DIR`.
@@ -29,6 +29,7 @@
 #' }
 cites_db_download <- function(tag = NULL, destdir = tempdir(),
                               cleanup = TRUE, verbose = interactive()) {
+
   if (verbose) message("Downloading data...\n")
   zfile <- get_gh_release_file("ropensci/citesdb",
     tag_name = tag,
@@ -39,14 +40,18 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
   temp_tsv <- tempfile(fileext = ".tsv")
   gunzip(zfile, destname = temp_tsv, overwrite = TRUE, remove = cleanup)
 
-  for (tab in dbListTables(cites_db())) {
-    dbRemoveTable(cites_db(), tab)
-  }
+  # for (tab in dbListTables(cites_db())) {
+  #   dbRemoveTable(cites_db(), tab)
+  # }
+  # cites_disconnect()
+  # invisible(cites_db())
+  unlink(cites_path(), recursive = TRUE, force = TRUE, expand = FALSE)
 
-  cites_disconnect()
   tblname <- "cites_shipments"
 
-  dbCreateTable(cites_db(), tblname, fields = cites_field_types)
+  try(dbRemoveTable(cites_db(read_only = FALSE), tblname), silent = TRUE)
+  cites_disconnect()
+  dbCreateTable(cites_db(read_only = FALSE), tblname, fields = cites_field_types)
 
   suppressMessages(
     dbExecute(
@@ -58,8 +63,8 @@ cites_db_download <- function(tag = NULL, destdir = tempdir(),
       )
     )
   )
-
-  dbWriteTable(cites_db(), "cites_status", make_status_table(version = ver),
+  cites_disconnect()
+  dbWriteTable(cites_db(read_only = FALSE), "cites_status", make_status_table(version = ver),
     overwrite = TRUE
[…]

### 2020-10-30 R/onattach.R

@@ -7,7 +7,6 @@ in_chk <- function() {
 }
 
 .onAttach <- function(libname, pkgname) {  #nolint
-  duckdb::duckdb_shutdown(duckdb::duckdb())
   if (interactive() && Sys.getenv("RSTUDIO") == "1"  && !in_chk()) {
     cites_pane()
   }
@@ -29,10 +28,8 @@ in_chk <- function() {
 #' }
 #' }
 cites_db_delete <- function() {
-  for (t in dbListTables(cites_db())) {
-    dbRemoveTable(cites_db(), t)
-  }
-  update_cites_pane()
+  cites_disconnect()
+  unlink(cites_path(), recursive = TRUE, force = TRUE, expand = FALSE)
 }
 
 
@@ -47,8 +44,8 @@ cites_db_delete <- function() {
 #' @examples
 #' cites_status()
 cites_status <- function(verbose = TRUE) {
-  if (dbExistsTable(cites_db(), "cites_shipments") &&
-      dbExistsTable(cites_db(), "cites_status")) {
+  if (DBI::dbExistsTable(cites_db(), "cites_shipments") &&
+      DBI::dbExistsTable(cites_db(), "cites_status")) {
     status <- DBI::dbReadTable(cites_db(), "cites_status")
     status_msg <-
       paste0(

### 2020-10-30 tests/testthat/test-02-tables.R

@@ -8,7 +8,7 @@ context("Tables")
 test_that("Tables have expected types", {
   skip_on_cran()
   skip_if_not(cites_status())
-  expect_is(cites_shipments(), "tbl_MonetDBEmbeddedConnection")
+  expect_is(cites_shipments(), "tbl_duckdb_connection")
   expect_is(cites_metadata(), "data.frame")
   expect_is(cites_codes(), "data.frame")
   expect_is(cites_parties(), "data.frame")

### 2020-10-30 tests/testthat/test-03-connection.R

@@ -8,16 +8,16 @@ context("Connection")
 test_that("Disconnetion works", {
   skip_on_cran()
   skip_if_not(cites_status())
-
   cites_disconnect()
   expect_error({
     success <- callr::r(function() {
-      options(CITES_DB_DIR = "localdb")
+      Sys.setenv(CITES_DB_DIR = normalizePath(file.path(getwd(), "localdb"),
+                                              mustWork = FALSE))
       con <- DBI::dbConnect(
         duckdb::duckdb(),
-        getOption("CITES_DB_DIR")
+        dbdir = file.path(Sys.getenv("CITES_DB_DIR"))
       )
-      out <- inherits(con, "MonetDBEmbeddedConnection")
+      out <- inherits(con, "duckdb_connection")
       citesdb::cites_disconnect()
       options(CITES_DB_DIR = NULL)
       return(out)
@@ -46,5 +46,5 @@ test_that("Tables fail when database is deleted", {
   expect_error(cites_parties())
 })
 
-unlink(Sys.getenv("CITES_DB_DIR"), recursive = TRUE)
+#unlink(Sys.getenv("CITES_DB_DIR"), recursive = TRUE)
 Sys.setenv(CITES_DB_DIR = olddir)
